### 2016-06-29 bioc-workflow.Rmd
Small updates to the text, typos etc

---
title: "A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data"
author: "Lisa Breckels and Laurent Gatto"
output: html_document
---

```{r env, echo=FALSE}
library("BiocStyle")
suppressPackageStartupMessages(library("knitr"))
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("pRoloc"))
suppressPackageStartupMessages(library("pRolocdata"))
```

# Introduction

Quantitative mass spectrometry based spatial proteomics involves
elaborate, expensive and time consuming experimental procedures and
considerable effort is invested in the generation of such
data. Multiple research groups have described a variety of approaches
to establish high quality proteome-wide datasets. However, data
analysis is as critical as data production for reliable and insightful
biological interpretation. Here, we walk the reader through a typical
pipeline for the analysis of such data using several Bioconductor
packages for the R statistical programming environment.

The main package to analyse protein localisation data is
`r Biocpkg("pRoloc")`, which offers a set of dedicated functions for the
analysis of such data. `r Biocpkg("pRoloc")` itself relies on `r Biocpkg("MSnbase")` to
manipulate and process quantitative proteomics data. Many other
packages are used by `r Biocpkg("pRoloc")` for clustering, classification and
visualisation.

In this workflow, we will describe how to prepare the spatial
proteomics data starting from a spreadsheet containing quantitative
mass spectrometry data. We will focus on a recent pluripotent mouse
embryonic stem cells experiment [@hyper]. Additional annotated and
pre-formatted datasets from various species are readily available in
the `r Biocexptpkg("pRolocdata")` package.

Installation of Bioconductor package is documented in details on the
[Bioconductor installation help page](http://bioconductor.org/install/#install-bioconductor-packages). This
procedure is also applicable to any packages, from
[CRAN](https://cran.r-project.org/) as well as GitHub. Once a package
has been installed, it needs to be loaded for it functionality become
available in the R session; this is done with the `library` function.

If you have questions about this workflow in particular, or about
other Bioconductor packages in general, they a best asked on the
[Bioconductor support site](https://support.bioconductor.org/)
following the
[posting guidelines](http://www.bioconductor.org/help/support/posting-guide/). Questions
can be tagged with specific package names or keywords. For more
general information about mass spectrometry and proteomics, the
readers are invited to read the `r Biocexptpkg("RforProteomics")`
package vignettes and associated papers [@Gatto:2014;@Gatto2015].

# Reading and processing spatial proteomics data

## The use-case: prediction sub-cellular localisation in pluripotent embryonic mouse stem cells

As a use-case, we analyse a recent high-throughput spatial proteomics
dataset from pluripotent mouse embryonic stem cells (E14TG2a)
[@hyper]. The data was generated using hyperplexed LOPIT (hyperLOPIT),
an improved method relying on improved sub-cellular fractionation and
more accurate quantitation, leading to more reliable classification of
protein localisation across the whole sub-cellular space. The method
uses an elaborate sub-cellular fractionation scheme, enabled by the
use of Tandem Mass Tag (TMT) [Thompson:2003] 10-plex and application
of the MS data acquisition technique named synchronous precursor
selection MS^3^ (SPS-MS^3^) [@McAlister:2014], for TMT
quantification with high accuracy and precision. Three biological
replicates were generated from the E14TG2a experiment, the first was
to target low density fractions and the second and third were to
emphasis separation of the denser organelles.  The intersect of
replicates 1 and 2 was treated as a 20-plex dataset for the analysis
discussed in the manuscript [@hyper] as it has been shown that
combining replicates across from different gradients can increase
spatial resolution [@Trotter:2010]. The combination of replicates resulted
in 5032 proteins common in both experiments.

These, as well as many other data are directly available as properly
structured and annotated computational object from the 
`r Biocexptpkg("pRolocdata")` experiment package. In this workflow, we
will start with a description of how to generate these ad hoc objects
starting from any arbitrary spreadsheets, as produced by many popular
third-party applications.

While we focus here on a LOPIT-type dataset, these analyses are
relevant for any quantitative spatial proteomics data, irrespective of
the fractionation or quantitation (i.e. labelled or label-free)
methods.

## The infrastructure: `r Biocpkg("pRoloc")` and `r Biocpkg("MSnbase")` in Bioconductor

To make use of the full functionality of the `r Biocpkg("pRoloc")`
software one needs to import their data into R and prepare them as an
`MSnSet`. The `MSnSet` is a dedicated data structure for the efficient
manipulation and processing of mass spectrometry and proteomics data
in R. Figure 1 illustrates a simplified view of the `MSnSet`
structure; there exists 3 key sub-parts (termed slots) to such a data
object: (1) the `exprs` slot for storing the quantitation data, (2)
the `fData` slot for storing the feature meta-data, and finally (3)
the `pData` slot for storing the sample meta-data.

![Simplified representation of the `MSnSet` data structure (reproduced with permission from the `r Biocpkg("MSnbase")` vignette)](./Figures/msnset.png)

## Importing data

There are a number of ways to import quantitation data and create an
`MSnSet` instance and all methods are described in the `r Biocpkg("MSnbase")`
[input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf). One
suggested simple method is to use the function `readMSnSet2` in 
`r Biocpkg("MSnbase")`. The function takes a single spreadsheet as input
and extracts the columns containing the quantitation data, as
identified by the argument `ecol`, to create the expression data,
while the other columns in the spreadsheet are appended to the feature
meta-data slot.  By example, in the code chunk below we read in the
`csv` spreadsheet containing the quantitation data from the intersect
of replicates 1 and 2 of the mouse map [@hyper], using the
`readMSnSet2` function. The data is as available online with the
manuscript (see tab 2 of the `xlsx` supplementary data set 1 in
[@hyper], which should be exported as a text-based spreadsheet) and
also as a `csv` in the Bioconductor `r Biocexptpkg("pRolocdata")` data
package.

To use the `readMSnSet2` function, as a minimum one must specify the
file path to the data and which columns of the spreadsheet contain
quantitation data. The `getEcols` function exists to help users
identify which columns of the spreadsheet contain the
quantitation data.

We first locate the the spreadsheet of E14TG2a data distributed with
the `r Biocexptpkg("pRolocdata")` data using `system.file`, that
locates the `extdata` in the `r Biocexptpkg("pRolocdata")` package on
the hard drive and `dir`, that displays the full path to the file
matching the `hyperLOPIT-SIData-ms3-rep12-intersect.csv` pattern,
which corresponds the the file of interest. In the last line, we print
the file name (not the full path, which will vary from computer to
computer).

```{r getFilename}
library("MSnbase")
extdatadir <- system.file("extdata", package = "pRolocdata")
csvfile <- dir(extdatadir, full.names = TRUE,
          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
basename(csvfile)
```

Note that the file is compressed (as indicated by the `gz`, for
`gzip`, extension), and will be decompressed on-the-fly when read into
R later on.

The spreadsheet that was deposited by the authors contains two
headers, with the second header containing information about where the
quantitation data is stored.

![A screenshot of the data in the spreadsheet.](./Figures/spreadsheet-screenshot.png)

We can display the names of the second header by calling the
`getEcols` function with the argument `n = 2` (the default value is `n
= 1`), to specify that we wish to display the column names of the
second line.

```{r getEcols}
getEcols(csvfile, split = ",", n = 2)
```

It is now easy for one to identify that the quantitation data,
corresponding to the 10 TMT isobaric tags, is located in columns 8
to 27. We now have the two mandatory arguments to `readMSnSet2`,
namely the file name (stored in the `csvfile` variable) and the
quantitation column indices. In addition to these, it is also possible
to pass the optional argument `fnames` to indicate which column to use
as the labels by which to identify each protein in the sample. Here,
we use `fnames = 1` to use the UniProt identifiers contained in the
first (unnamed) column of the spreadsheet. We also need to specify to
skip the first line of the file (for the same reason that we used 
`n = 2` in `getEcols` above) to read the `csv` data and convert it to an
`MSnSet` object, named `hl` (for hyperLOPIT).

```{r readMSnSet2}
hl <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1)
```

Below, we display a short summary of the data. The data contains 
`r nrow(hl)` proteins/features common across the 2 biological replicates
for the respective 2 x 10-plex reporter tags (`r ncol(hl)`
columns/samples), along with associated feature meta-data such as
protein markers, protein description, number of quantified peptides
etc (see below).

```{r showhl}
hl
```

As briefly mentioned above, the quantitation data is stored in the
`exprs` slot of the `MSnSet` and can be accessed by
`exprs(hl)`. Below, we examine the quantitative information for
first 5 proteins. It is also possible to access specific rows and
columns by naming the proteins and fractions of interest.

```{r assesshl}
exprs(hl)[1:5, ]
exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
```

The feature meta-data is stored in the `fData`
slot and can be accessed by `fData(hl)`. When using
`readMSnSet2`, automatically, everything that is not defined as
quantitation data by `ecol` or the feature names by `fnames` is
deposited to the `fData` slot. As we wish to demonstrate the complete
analysis of this data we remove the results from the prior analysis
described in [@hyper] in the code chunk below. We see the `fData`
contains 25 columns describing information such as the number of
peptides, associated markers, machine learning results etc.

For demonstration in the code chunk below keep the 2nd, 8th and 11th
columns which contain the UniProt entry names and two different marker
sets to use an input for machine learning analyses (see sections on
markers and subsequent sections) and rename the first feature variable
names.

```{r removefData}
fvarLabels(hl)
fData(hl) <- fData(hl)[, c(2, 8, 11)]
fvarLabels(hl)[1] <- "UniProtId"
head(fData(hl))
```

Note that when using the simple `readMSnSet2` procedure, sample
meta-data is kept empty. It is advised to annotate the fractions as
well. Below, we annotate the replicate from which the profiles
originate and the TMT tag (extracted from the sample/fraction names).

```{r pdata}
pData(hl)$replicate <- rep(1:2, each = 10)
pData(hl)$tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
pData(hl)
```

## Data processing

### Normalisation

There are two aspects related to data normalisation that are relevant
to spatial proteomics data processing. The first one focuses on
reducing purely technical variation between fractions without
affecting biological variability (i.e. the shape of the quantitative
profiles) too much, as generally focused on. This normalisation will
depend on the underlying quantitative technology and the experimental
design, and will not be addressed in this workflow. In addition, and
more specific to spatial proteomics data, it is fundamental to scale
all the organelle-specific profiles into a same intensity interval
(typically 0 and 1) by, for example, diving each intensity by the sum
of the intensities for that quantitative feature, as illustrated
below.

```{r normsum, eval = FALSE}
hl <- normalise(hl, method = "sum")
```

This transformation of the data assures to cancel the effect of the
absolute intensities of the quantitative features along the rows, to
focus subsequent analyses on the relative profiles along the
sub-cellular fractions.

The same `normalise` function (or `normalize`, both spellings are
supported) can also be applied in the first case described
above. Different normalisation methods such as mean or median scaling,
variance stabilisation or quantile normalisation, to cite a few, can
be applied to accomodation different needs.

Before combination, the two replicates in the `hl` data that we
read into R were separately normalised by sum across the 10 channels
for each replicate respectively. We can verify this by summing each
rows for each replicate:

```{r normcheck}
summary(rowSums(exprs(hl[, hl$replicate == 1])))
summary(rowSums(exprs(hl[, hl$replicate == 2])))
```
We see that some features do not add up exactly to 1 due to rounding
errors after exporting to intermediate files. These small deviations
do not bear any consequences here.

### Combining acquisitions

The spreadsheet that was used to create the `hl` MSnSet
included two replicates. We also provide individual replicates in the
`r Biocexptpkg("pRolocdata")` package. Below, we show how to combine
`MSnSet` objects and, subsequently, how to filter and handle missing
values. We start by loading the experimental data package and the
equivalent replicates using the `load` function.

```{r reps4combine}
library("pRolocdata")
data(hyperLOPIT2015ms3r1)
data(hyperLOPIT2015ms3r2)
```

To list all data distributed with a package, one can use pass the
package name as input to the `data` function. For instance

```{r pkgdata, eval=FALSE}
data(package = "pRolocdata")
```

will list the `r nrow(data(package = "pRolocdata")$results)` datasets
that are available in `r Biocexptpkg("pRolocdata")`.

Combining data is performed with the `combine` function. This function
will inspect the feature and sample names to identify how to combine
the data. As we want our replicates to be combined along the columns
(same proteins, different sets of fractions), we need to assure that
the respective sample names differ so they can be identified from one
another. The function `updateSampleNames` can be used do this.

```{r samplenames}
identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
sampleNames(hyperLOPIT2015ms3r1)
sampleNames(hyperLOPIT2015ms3r2)
```

In addition, to matching names, the content of the feature metadata
for identical feature annotations must match exactly across the data
to be combined. In particular for these data, we expect the same
proteins in each replicates to be annotated with the same UniProt
entry names and descriptions, but not with the same coverage of number
of peptides or peptide-spectrum matches (PSMs).

```{r fvarnames}
fvarLabels(hyperLOPIT2015ms3r1)
fvarLabels(hyperLOPIT2015ms3r2)
```

Below, we update the replicate specific feature variable names and
remove the shared annotation.

```{r fvarnames2}
fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
fvarLabels(hyperLOPIT2015ms3r2)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r2)[3:5], 2)
fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
```

We can now combine the two experiments into a single `MSnSet`:

```{r combine}
hyperLOPIT <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
hyperLOPIT
```

More details above combining data are given in the dedicated
*Combining MSnSet instances* section of the `r Biocpkg("MSnbase")`
[tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).

### Missing data

Missing data are a recurrent issue in mass spectrometry applications,
and should be addressed independently of this workflow
[@Webb-Robertson:2015;@Lazar:2016]. In [@Gatto:2014b], we have
described how a high content in missing values in spatial proteomics
data and their inappropriate handling leads to a reduction of
sub-cellular resolution. Missing data can be imputated using

`r Biocpkg("MSnbase")`'s `impute` function. The method underlying the
imputation method is then determined by a `methods` parameter. In our
particular case, missing values are indicative of protein groups that
were not profiles in both replicates.

```{r namap}
image2(is.na(hyperLOPIT), col = c("black", "white"),
       main = "Missing values (white cells) after combining replicates")
```

We prefer to remove proteins that were not assayed in both of our two
replicated experiments. This is done with the `filterNA` function that
removes features that contain more than a certain proportion (default
is 0) missing values.

```{r filterNA}
hyperLOPIT <- filterNA(hyperLOPIT)
hyperLOPIT
```

When more than 2 data are to be combined and/or too many proteins were
not consistently assayed, leading to too many proteins being filtered
out, we suggest to implement an ensemble of classifiers voting on
protein-sub-cellular niche membership over the output of several
experiments (see section *Supervised machine learning* for the
description of sub-cellular assignments).

# Quality Control

Data quality is routinely examined through visualisation to verify
that sub-cellular niches have been separated along the gradient. Based
on De Duve's principle [@DeDuve:1981] proteins that co-localise
exhibit similar quantitation profiles across the gradient fractions
employed. One approach that has been widely used to visualise and
inspect high throughput mass spectrometry-based proteomics data is
principal components analysis (PCA). PCA is one of many dimensionality
reduction methods, that allow one to effectively summarise
multi-dimensional data in to 2 or 3 dimensions to enable
visualisation. Very generally, the original continuous
multi-dimensional data is transformed into a set of orthogonal
components ordered according to the amount of variability that they
describe. The `plot2D` method in `r Biocpkg("pRoloc")` allows one to
plot the principal components (PCs) of a dataset against one another,
by default the first two components are plotted on the x- and y-axis,
respectively (the `dims` argument can be used to plot other PCs). If
distinct clusters are observed, we assume that there is organellar
separation present in the data. In the code chunk below we produce a
PCA plot of the mouse stem cell dataset. One point on the plot
represents one protein. We can indeed see several distinct protein
clusters. Although, representing the multi-dimensional data along a
limited set of PCs does not give us a hard quantitative measure of
separation, it is extremely useful summarising complex experimental
information in one figure, to get an simplified overview of the data.

In the code chunk below, we specify `fcol = NULL`, which means not to
consider any feature variable to annotate the features (proteins) with
colours. We will see later how to use this to annotate the PCA plot
with prior information about sub-cellular localisation.

```{r qcplot, message=FALSE, warning=FALSE}
library("pRoloc")
plot2D(hl, fcol = NULL)
```

In the first instance we advise one to visualise their data without 
any annotation (i.e. with `fcol = NULL`), before proceeding with data 
annotation. The identification of well resolved clusters in the data, 
constitutes an unbiased assessment of the data structure, demonstrating 
the successful separation of sub-cellular clusters.

**TODO** There are other dimensionality reduction methods available in
the `plot2D` function, which can be parametrised with the `method`
argument.  Advantages, disadvantages? See
[this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
very interesting gist with intriguing tSNE examples. This might be
better to come after markers.

It is also useful to visualise the relative intensities along the
gradient to identify fractions displaying particularly low yield. This
can be done using the `plotDist` and `boxplot` functions, that plot
the protein profiles occupancy along the gradient (we also display the
mean fraction intensities) and a `boxplot` of the column
intensities. In the two plots displayed below, we re-order the
fractions to pair corresponding fractions in the two replicates
(rather than ordering the fractions by replicate).

```{r qcbx, fig.width = 12}
par(mfrow = c(1, 2))
o <- order(hyperLOPIT$Fraction.No)
plotDist(hyperLOPIT[, o], pcol = "#00000010")
lines(colMeans(exprs(hyperLOPIT[, o])), col = "red", type = "b")
boxplot(exprs(hyperLOPIT[, o]))
```

# Markers

In the context of spatial proteomics, a marker protein is defined as a
well-known resident of a specific sub-cellular niche in a species
*and* condition of interest. Applying this to machine learning (ML),
and specifically supervised learning, for the task of protein
localisation prediction, these markers constitute the labelled training
data to use as input to a classification analyses. Defining well-known
residents, and obtaining labelled training data for ML analyses can be
time consuming, but it is important to define markers that are
representative of the multivariate data space and on which a
classifier will be trained and generated. `r Biocpkg("pRoloc")`
provides a convenience function, `addMarkers`, to directly add markers
to a `MSnSet` object, as demonstrated in the code chunk below. These
marker sets can be accessed using the `pRolocmarkers()`
function. Marker sets are stored as a simple named vector in R, and
originate from in-house user-defined sets of markers or from previous
published studies [@Gatto:2014b], which are continuosly updated and 
integrated.

```{r markers}
## List available marker sets
pRolocmarkers()
```

These markers can then be mapped to an `MSnSet`'s `featureNames`. The
mouse dataset used here has Uniprot IDs stored as the `featureNames`
(see `head(featureNames(lopit2016))`) and the names of the vector of
the mouse markers stored in `r Biocpkg("pRoloc")` (`mmus` markers) 
are also Uniprot IDs (see `head(mrk)` in the code chunk below), so 
it is straightforward to match names between the markers and 
the `MSnSet` instance using the `addMarkers` function.

```{r addmrkers}
## Use mouse markers
mrk <- pRolocmarkers(species = "mmus")
head(mrk)

## Add mouse markers
hl <- addMarkers(hl, mrk)
```

If the naming between the marker sets and the `MSnSet` dataset are
different, one will have to convert and match the proteins according
to the appropriate identifier. Sometimes, we find the equivalent entry
name, Uniprot ID or accession number is stored with the data, which
makes conversion between identifers relatively straightforward. If
this is not the case however, conversion can be performed using the 
`r Biocpkg("biomaRt")`, the Bioconductor
[annotation resouces](http://bioconductor.org/help/workflows/annotation/Annotation_Resources/)
or any conversion softwares available online.

We now visualise these annotations along the PCA plot using the
`plot2D` function. We also use the `addLegend` function to map the
marker classes to the pre-defined colours. We also display the data
along the first and seventh PCs using the `dims` argument. Note that
in this second call to the `plot2D` function, we have omitted the
`fcol` argument, as the `"markers"` feature variable name is the
default value. We choose to display PCs 1 and 7 to illustrate that
while upper principal components explain much less variability in the
data (2.23% for PC7, as opposed to 48.41% for PC1), we see that the
mitochondrial and peroxisome clusters can be differenciated, despite
the apparent overlap in the two first PCs.

```{r plotmarkers, fig.width = 12}
par(mfrow = c(1, 2))
plot2D(hl, fcol = "markers", main = "pRolocmarkers for mouse")
addLegend(hl, fcol = "markers", where = "bottomleft", cex = .7)
plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
```

The colours have been defined so as to enable to differenciate up to
30 classes. If more are provided, different character symbols
(circles, squares, ... and empty and solid symbols) are used. The
colours and the default plotting characters (solid dots for the
markers and empty circles for the features of unknown localisation)
can of course be changed, as described in the `setStockcol` manual
page.

In general, the Gene Ontology (GO) [@Ashburner:2000], and in
particular the cellular compartment (CC) namespace are a good starting
point for protein annotation and marker definition. It is important to
note however that automatic retrieval of sub-cellular localisation
information, from `r Biocpkg("pRoloc")` or elsewhere, is only the
beginning in defining a marker set for downstream analyses. Expert
curation is vital to check that any annotation added is in the correct
context for the the biological question under investigation.

Another useful visualisation that relies on marker annotation is the
representation of the protein profiles occupancy along the gradient
using the `plotDist` function. While the PCA plot enables to
efficiently visualise the complete dataset and assess the relative
separation of different sub-cellular niches, comparing profiles of a
few marker clusters is useful to assess how exactly they differ (in
terms of peak fractions, for example). Below, we plot the profile of
the mitochondrial and peroxisome markers to highlight the differences
in profiles between these two sets of markers along the 6th and 7th
fractions, as represented above along the 7th PC on the PCA plot.

```{r plotDist}
hlo <- hyperLOPIT[, order(hyperLOPIT$Fraction.No)]
plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
         pcol = "purple", fractions = "Fraction.No")
title(main = "Marker occupancy profiles along the gradient")
matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
         lty = 1, col = "darkblue", type = "l")
legend("topleft", c("Mitochondrion", "Peroxisome"),
       lty = 1, col = c("purple", "blue"), bty = "n")
```

# Replication

With the aim of maximising the sub-cellular resolution and,
consequently, the reliability in protein sub-cellular assignments, we
follow the advice in [@Trotter:2010] and combine replicated spatial
proteomics experiments as described above. Indeed, Trotter et
al. have shown a significant improvement in protein–organelle
association upon direct combination of single experiments, in
particular when these resolve different subcellular niches.

Direct comparisons of individual fractions in replicated experiments
does not provide an adequate, goal-driven assessment of different
experiments. Indeed, due to the nature of the experiment and fraction
collection, the quantitative channels (loosely referred to as
fractions) do not correspond to identical actual fractions along the
gradient. As can be seen in the table below (taken from the
`hyperLOPIT`'s `pData`, focusing on channels 7 to 10), different sets
of fractions are combined to obtain enough material and optimise
acurate quantitation.

```{r pdtab, echo=FALSE}
i <- hyperLOPIT$Fraction.No %in% 7:10
pData(hyperLOPIT)[i, -3]
```

<!-- On the figure below, we compare the relative intensities of channels 7 -->
<!-- to 10 on scatter plots, highlighting different marker sets. These -->
<!-- differences result from different peak fractions for these niches on -->
<!-- the mitochondrial and chromatin occupancy profiles for the two -->
<!-- replicates. -->

<!-- ```{r repl1, eval = TRUE, echo=FALSE, fig.width = 12} -->
<!-- sel0 <- fData(hyperLOPIT)$markers != "unknown" -->
<!-- cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)] -->
<!-- cls[!sel0] <- getUnknowncol() -->

<!-- par(mfrow = c(2, 4)) -->
<!-- for (i in c(7, 8, 10, 9)) { -->
<!--     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls, -->
<!--          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"])) -->
<!--     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19) -->
<!--     grid() -->
<!--     abline(0, 1, col = "red") -->
<!-- } -->
<!-- plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "") -->
<!-- addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7) -->

<!-- i <- fData(hyperLOPIT)$markers == "Mitochondrion" -->
<!-- plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2], -->
<!--          featureNames(hyperLOPIT)[i], type = "l", -->
<!--          mcol = "orange") -->
<!-- title(main = "Mitochondrion") -->
<!-- ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2]) -->
<!-- matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1) -->
<!-- legend("topleft", paste("Replicate", 1:2), -->
<!--        col = c("#F781BF", "orange"), -->
<!--        lty = 1, bty = "n") -->

<!-- i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin" -->
<!-- plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2], -->
<!--          featureNames(hyperLOPIT)[i], type = "l") -->
<!-- ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2]) -->
<!-- matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1) -->
<!-- title(main = "Chromatin") -->
<!-- legend("topleft", paste("Replicate", 1:2), -->
<!--        col = c("#984EA3", "steelblue"), -->
<!--        lty = 1, bty = "n") -->
<!-- ``` -->

The more relevant comparison unit is not a single fraction, but rather
the complete protein occupancy profiles, which are best visualised as
experiment-wide on a PCA plot. As such, we prefer to focus on the
direct, qualitative comparison of individual replicate PCA plots,
assuring that each displays acceptable sub-cellular resolution. Note
that in the code chunk below, we mirror the x-axis to represent the
two figures with the same orientation.

```{r plot2Drep, fig.width = 12}
par(mfrow = c(1, 2))
plot2D(hl[, hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
plot2D(hl[, hl$replicate == 2], fcol = "SVM.marker.set", main = "Replicate 2",
       mirrorX = TRUE)
```

Peason correlation

```{r}
i1 <- exprs(hl[, hl$replicate == 1])
i2 <- exprs(hl[, hl$replicate == 2])
pearcor <- cor(t(i1), t(i2))
pearcor0 <- cor(t(i1[, seq(1, 10, 2)]), t(i2[, seq(1, 10, 2)]))
mean(diag(pearcor))
mean(diag(pearcor0))
```

- What about classification on different replicates?

# Interactive visualisation

A section about  `r Biocpkg("pRolocGUI")`.

# Novelty Detection

The extraction of sub-cellular protein clusters can be difficult owing to the limited number of marker proteins that exist in databases and elsewhere. Furthermore, given the vast complexity of the cell, automatic annotation retrieval does not always give a full representation of the true sub-cellular diversity in the data. For downstream analyses, such as supervised machine learning, it is desirable to obtain reliable markers that cover as many sub-cellular niches as possible, as these markers are directly used in the training phase of the ML classification. We find that a lack of sub-cellular diversity in the labelled training data leads to prediction errors, as unlabelled instances can only be assigned to a class that exists in the training data [@Breckels:2013]. In such scenarios novelty detection can be useful to identify data-specific sub-cellular groupings such as organelles and protein complexes. The phenotype discovery (phenoDisco) algorithm [@Breckels:2013] is one such method and is available in `r Biocpkg("pRoloc")`. It is an iterative semi-supervised learning method that combines the classification of proteins on existing labelled data with the detection of new clusters. Novelty detection methods are also useful for confirming the presence of known clusters in an unbiased fashion. For example, in [@hyper] the `phenoDisco` algorithm was used to detect the data-specific confirmation and presence of the nucleus and nucleus sub-compartments, as it was expected that nucleus associated clusters existed within the data as the protocol included a separate chromatin-rich fraction to enrich for nuclei. In the code chunk below, we show how to run a series of novelty detection experiments using the `phenoDisco` algorithm on the mouse stem-cell dataset. There are several optional arguments that can be set dependent on the type of biological question of interest. For example,

- Explain how to use different group sizes for detecting small complexes etc.
- ndims for using >=2 components
- times -> important for convergence
- input markers are different here -> explain choice of pd.markers

```{r pd, eval = FALSE}
## As per the hyperLOPIT paper
pdRes <- phenoDisco(hl, fcol = "phenoDisco.Input", times = 200, 
                    GS = 20, p = 0.05)

```

# Supervised machine learning

Supervised machine learning, also known as classification, is an essential tool for the assignment of proteins to distinct sub-cellular niches. Using a set of labelled training examples i.e. markers, we can train a machine learning classifier to learn a mapping between the data i.e. the quantitative protein profiles, and a known localisation. The trained classifier can then be used to predict the localisation of a protein of unknown localisation, based on its observed protein profile. To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @Trotter:2010; @Hall:2009; @Dunkley:2006; @Tan:2009].

There are several classification algorithms readily available in `pRoloc`, which are documented in the dedicated [`pRoloc` machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf). We find the general tendancy to be that it is not the choice of classifier, but the improper optimisation of the algorithmic parameters, that limits classification accuracy. Before employing any classification algorithm and generating a model on the training data, one must first find the optimal parameters for the algorithm of choice.

## Optimisation

In the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data. The training data is found in the `featureData` slot in the column called `SVM.marker.set`. As previously mentioned one first needs to train the classifiers parameters before an algorithm can be used to predict the class labels of the proteins with unknown location. One of the most common ways to optimise the parameters of a classifier is to partition the labelled data in to training and testing subsets. In this framework parameters are tested via a grid search using cross-validation on the training partition. The best parameters chosen from the cross-validation stage are then used to build a classifier to predict the class labels of the protein profiles on the test partition. Observed and expected classication results can be compared, and then used to assess how well a given model works by getting an estimate of the classiers ability to achieve a good generalisation i.e. that is given an unknown example predict its class label with high accuracy. In `pRoloc` algorithmic performance is estimated using stratified 80/20 partitioning for the training/testing subsets respectively, in conjuction with five-fold cross-validation in order to optimise the free parameters via a grid search. This procedure is usually repeated 100 times and then the best parameter(s) are selected upon investigation of classifier accuracy, here we use the harmonic mean of precision and recall; the macro F1 score. In the code chunk below we demonstrate how to optimise the free parameters; `sigma` and `cost`, of a classical SVM classifier with a Gaussian kernel using the function `svmOptimisation`. As the number of labelled instances per class varies from organelle to organelle, we can account for class imbalance by setting specific class weights when generating the SVM model. Below the weights, `w` are set to be inversely proportional to the class frequencies.

```{r loadpdRes, include=FALSE}
pdRes <- hl
load("Data/fusionparams.rda")
(w <- table(fData(pdRes)[, "SVM.marker.set"]))
w <- 1/w[names(w) != "unknown"]
```

```{r optimise, eval=FALSE}
(w <- table(fData(pdRes)[, "SVM.marker.set"]))
w <- 1/w[names(w) != "unknown"]

## 100 rounds of optimisation with five-fold cross-validation
params <- svmOptimisation(pdRes, fcol = "SVM.marker.set",
                          times = 100, xval = 5,
                          class.weights = w,
                          verbose = FALSE)
```

The output `params` is an object of class `GenRegRes`; a dedicated container for the storage of the design and results from a machine learning optimisation. To assess classifier performance we can examine the macro F1 scores and the most frequently chosen parameters. A high macro F1 score indicates that the marker proteins in the test dataset are consistently and correctly assigned by the the algorithm. Often more than one parameter or set of parameters gives rise to the best generalisation accuracy. As such it is always important to investigate the model parameters and critically assess the best choice. The best choice may not be as simple as the parameter set that gives rise to the highest macro F1 score and one must be careful to avoid overfitting and to choose parameters wisely.

```{r visualiseOpt, eval=FALSE}
(best <- getParams(params))
plot(params)
levelPlot(params)
```
```{r visualiseOptHide, echo=FALSE, fig.width = 12}
grid.arrange(plot(params), levelPlot(params), ncol = 2)
```

By using the function `getParams` we can extract the best set of parameters. Currently, `getParams` retrieves the first best set is automatically but users are encouraged to critically assess whether this is the most wise choice by visualising the results using the methods `plot` and `levelPlot`. The `plot` method for `GenRegRes` object shows the respective distributions of the 100 macro F1 scores for the best cost/sigma parameter pairs, and `levelPlot` shows the averaged macro F1 scores, for the full range of parameter values. Once we have selected the best parameters we can then use them to build a classifier from the labelled marker proteins.

## Classification

We can use the function `svmClassification` to return a classification result for all unlabelled instances in the dataset corresponding to their most likely sub-cellular location. The algorithm parameters are passed to the function, along with the class weights and the `fcol` to tell the function where the labelled training data is located, in our case, our marker proteins are located in the `featureData` with the label `"SVM.marker.set"`.

```{r classify, eval=TRUE}
svmRes <- svmClassification(pdRes, params,
                            class.weights = w,
                            fcol = "SVM.marker.set")
```

Automatically, the output of the above classification; the organelle predictions and assignment scores, are stored in the `featureData` slot of the `MSnSet`. In this case, they are given the labels `svm` and `svm.scores` for the predictions and scores respectively. The resultant predictions can be visualised using `plot2D`. In the code chunk below `plot2D` is called to generate a PCA plot of the data and `fcol` is used to specify where the new assignments are located e.g. `fcol =  "svm"`.

Additionally, when calling `plot2D` we can use the `cex` argument to change the size of each point on the plot (where one point represents one protein) to be inversely proportional to the SVM score. This gives an initial overview of the high scoring localisations from the SVM predictions.

```{r setColors, include=FALSE}
setStockcol(paste0(getStockcol(), 70))
```
```{r plotSVM, eval=TRUE}
## set point size of each protein to be inversely proportional to the SVM score
ptsze <- exp(fData(svmRes)$svm.scores) - 1

## plot new predictions
plot2D(svmRes, fcol = "svm", cex = ptsze)
addLegend(svmRes, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
```

The adjustment of the point size intuitively confers important
information that is more difficult to defined formally, but that we
will address in the next section. The classifier (SVM in our case, but
this is also valid of other classifiers) defines boundaries based on
the labelled marker proteins. These class/organelle boundaries define
how non-assigned proteins are classifier and with what
confidence. **TODO** explain how choosing markers
positively/negatively affects this, and relation with thresholding.

## Thresholding

It is common when applying a supervised classification algorithm to set a specific score cutoff on which to define new assignments, below which classifications are set to unknown/unassigned. This is because in a supervised learning setup proteins can only be predicted to be localised to one of the sub-cellular niches that appear in the labelled training data. We can not guaranette (and do not expect) that the whole class diversity to be represented in the training data as (1) finding markers that represent the whole diversity of the cell is challenging and (2) many sub-cellular niches contain too few proteins to train on (note: we recommend a minimum of 13 markers per sub-cellular class for stratified 80/20 partitioning and cross-validation).

Deciding on a threshold is not trivial as classifier scores are heavily dependent upon the classifier used and different sub-cellular niches can exhibit different score distributions. We recommend one to set class-specific score distributions. In the code chunk below we display a boxplot of the score distributions per organelle.

```{r thresholds}

```

To help examine these distributions there is also a dedicated interactive application in the `r Biocpkg("pRolocGUI")` package. To access this application:

```{r classifyApp}

```

To launch the classify application:

- Essential for protein localisation as we do not want to force all proteins in to one category (allow for multiple lcoalisations or localisations outside the training data set e.g. new novel compartments)

- `getPredictions` function allows you to pass a score threshold. Also `classify` application available to interactively view output of thresholds

## Transfer learning

# Session information

The function `sessionInfo` provides a summary of all packages and
versions used to generate this document. This enables us to record the
exact state of our session that lead to these exact
results. Conversely, if the script stops working of if it returns
different results, we are in a position to re-generate the original
results using the adequate software versions and retrace changes in
the software that lead to failure and/or different results.

```{r si}
sessionInfo()
```

It is always important to include session information details along
with a
[short reproducible example](http://adv-r.had.co.nz/Reproducibility.html)
highlighting the problem or
[question](https://support.bioconductor.org/) at hand.