### 2016-09-29 bioc-workflow.Rnw
move to Rnw, update license to CC-BY

<<env, echo=FALSE, warning=FALSE>>=
library("BiocStyle")
suppressPackageStartupMessages(library("knitr"))
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("pRoloc"))
suppressPackageStartupMessages(library("pRolocdata"))
cache <- TRUE
@

\section*{Introduction}

Quantitative mass spectrometry based spatial proteomics involves
elaborate, expensive and time consuming experimental procedures and
considerable effort is invested in the generation of such data.
Multiple research groups have described a variety of approaches to
establish high quality proteome-wide datasets. However, data analysis
is as critical as data production for reliable and insightful
biological interpretation. Here, we walk the reader through a typical
pipeline for the analysis of such data using several Bioconductor
packages for the R statistical programming environment.

The main package to analyse protein localisation data is
\Biocpkg{pRoloc}, which offers a set of dedicated functions for the
analysis of such data. \Biocpkg{pRoloc} itself relies on
\Biocpkg{MSnbase} to manipulate and process quantitative
proteomics data. Many other packages are used by \Biocpkg{pRoloc}
for clustering, classification and visualisation.

In this workflow, we will describe how to prepare the spatial
proteomics data starting from a spreadsheet containing quantitative
mass spectrometry data. We will focus on a recent pluripotent mouse
embryonic stem cells experiment \cite{hyper}. Additional annotated and
pre-formatted datasets from various species are readily available in
the \Biocexptpkg{pRolocdata} package.

Installation of Bioconductor package is documented in details on the
\href{http://bioconductor.org/install/#install-bioconductor-packages}{Bioconductor
  installation help page}. This procedure is also applicable to any
packages, from \href{https://cran.r-project.org/}{CRAN} as well as
GitHub. Once a package has been installed, it needs to be loaded for
its functionality to become available in the R session; this is done
with the \texttt{library} function e.g.  to load the
\Biocpkg{pRoloc} one would type \texttt{library("pRoloc")}
after installation.

If you have questions about this workflow in particular, or about
other Bioconductor packages in general, they are best asked on the
\href{https://support.bioconductor.org/}{Bioconductor support site}
following the
\href{http://www.bioconductor.org/help/support/posting-guide/}{posting
  guidelines}. Questions can be tagged with specific package names or
keywords. For more general information about mass spectrometry and
proteomics, the readers are invited to read the
\Biocexptpkg{"RforProteomics"} package vignettes and associated
papers \cite{Gatto:2014,Gatto:2015}.

\section*{Reading and processing spatial proteomics data}

\subsection*{The use-case: predicting sub-cellular localisation in pluripotent embryonic mouse stem cells}

As a use-case, we analyse a recent high-throughput spatial proteomics
dataset from pluripotent mouse embryonic stem cells (E14TG2a)
\cite{hyper}. The data was generated using hyperplexed LOPIT
(hyperLOPIT), a state-of-the-art method relying on improved
sub-cellular fractionation and more accurate quantitation, leading to
more reliable classification of protein localisation across the whole
sub-cellular space. The method uses an elaborate sub-cellular
fractionation scheme, enabled by the use of Tandem Mass Tag (TMT)
\cite{Thompson:2003} 10-plex and application of the MS data
acquisition technique named synchronous precursor selection MS$^3$
(SPS-MS$^3$) \cite{McAlister:2014}, for TMT quantification with high
accuracy and precision. Three biological replicates were generated
from the E14TG2a experiment, the first was to target low density
fractions and the second and third were to emphasis separation of the
denser organelles.  The intersect of replicates 1 and 2 was treated as
a 20-plex dataset for the analysis.  As discussed in the manuscript
\cite{hyper} as it has been shown that combining replicates across
from different gradients can increase spatial resolution
\cite{Trotter:2010}. The combination of replicates resulted in 5032
proteins common in both experiments.

These, as well as many other data are directly available as properly
structured and annotated computational object from the 
\Biocexptpkg{pRolocdata} experiment package. In this workflow, we
will start with a description of how to generate these ad hoc objects
starting from an arbitrary spreadsheets, as produced by many popular
third-party applications.

While we focus here on a LOPIT-type dataset, these analyses are
relevant for any quantitative spatial proteomics data, irrespective of
the fractionation or quantitation (i.e. labelled or label-free)
methods.

\subsection*{The infrastructure: \texttt{pRoloc} and \texttt{MSnbase} packages}

To make use of the full functionality of the \Biocpkg{pRoloc}
software one needs to import their data into R and prepare them as an
\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for the efficient
manipulation and processing of mass spectrometry and proteomics data
in R. Figure 1 illustrates a simplified view of the \texttt{MSnSet}
structure; there exists 3 key sub-parts (termed slots) to such a data
object: (1) the \texttt{exprs} slot for storing the quantitation data, (2)
the \texttt{fData} slot for storing the feature meta-data, and finally (3)
the \texttt{pData} slot for storing the sample meta-data.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=.45\textwidth]{./Figures/msnset.png}
  \caption{Simplified representation of the \texttt{MSnSet} data
    structure (reproduced with permission from the \Biocpkg{MSnbase}
    vignette)}
  \label{fig:msnset}
\end{figure}

\subsection*{Importing data}

There are a number of ways to import quantitation data and create an
\texttt{MSnSet} instance. All methods are described in the
\Biocpkg{MSnbase}
\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf}{input/output
  capabilities vignette}. One suggested simple method is to use the
function \texttt{readMSnSet2} in \Biocpkg{MSnbase}. The function
takes a single spreadsheet as input and extracts the columns
containing the quantitation data, as identified by the argument
\texttt{ecol}, to create the expression data, while the other columns
in the spreadsheet are appended to the feature meta-data slot.  By
example, in the code chunk below we read in the \texttt{csv}
spreadsheet containing the quantitation data from the intersect of
replicates 1 and 2 of the mouse map \cite{hyper}, using the
\texttt{readMSnSet2} function. The data is as available online with
the manuscript (see tab 2 of the \texttt{xlsx} supplementary data set
1 in \cite{hyper}, which should be exported as a text-based
spreadsheet). It is also available as a \texttt{csv} in the
Bioconductor \Biocexptpkg{pRolocdata} data package.

To use the \texttt{readMSnSet2} function, as a minimum one must specify the
file path to the data and which columns of the spreadsheet contain
quantitation data. The \texttt{getEcols} function exists to help users
identify which columns of the spreadsheet contain the
quantitation data. In the last line of the code chunk below, we print
the file name (not the full path, which will vary from computer to
computer).

<<getFilename, cache = cache>>=
library("MSnbase")
extdatadir <- system.file("extdata", package = "pRolocdata")
csvfile <- dir(extdatadir, full.names = TRUE,
          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
basename(csvfile)
@

Note that the file is compressed (as indicated by the \texttt{gz}, for
\texttt{gzip}, extension), and will be decompressed on-the-fly when read into
R later on.

The spreadsheet that was deposited by the authors contains two
headers, with the second header containing information about where the
quantitation data is stored.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=.65\textwidth]{./Figures/spreadsheet-screenshot.png}
  \caption{A screenshot of the data in the spreadsheet.}
  \label{fig:spreadsheet}
\end{figure}

We can display the names of the second header by calling the
\texttt{getEcols} function with the argument \texttt{n = 2} (the default value is \texttt{n
= 1}), to specify that we wish to display the column names of the
second line.

<<getEcols>>=
getEcols(csvfile, split = ",", n = 2)
@

It is now easy for one to identify that the quantitation data,
corresponding to the 10 TMT isobaric tags, is located in columns 8
to 27. We now have the two mandatory arguments to \texttt{readMSnSet2},
namely the file name (stored in the \texttt{csvfile} variable) and the
quantitation column indices. In addition to these, it is also possible
to pass the optional argument \texttt{fnames} to indicate which column to use
as the labels by which to identify each protein in the sample. Here,
we use \texttt{fnames = 1} to use the UniProt identifiers contained in the
first (unnamed) column of the spreadsheet. We also need to specify to
skip the first line of the file (for the same reason that we used 
\texttt{n = 2} in \texttt{getEcols} above) to read the \texttt{csv} data and convert it to an
\texttt{MSnSet} object, named \texttt{hl} (for hyperLOPIT).

<<readMSnSet2>>=
hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
@

Below, we display a short summary of the data. The data contains 
\Sexpr{nrow(hl)} proteins/features common across the 2 biological replicates
for the respective 2 x 10-plex reporter tags (\Sexpr{ncol(hl)}
columns/samples), along with associated feature meta-data such as
protein markers, protein description, number of quantified peptides
etc (see below).

<<showhl>>=
hl
@

Below, we examine the quantitative information for first 5 proteins.
It is also possible to access specific rows and columns by naming the
proteins and TMT tag channels of interest.

<<assesshl>>=
exprs(hl)[1:5, ]
exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
@

The feature meta-data is stored in the \texttt{fData} slot and can be
accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2}, automatically,
everything that is not defined as quantitation data by \texttt{ecol} or the
feature names by \texttt{fnames} is deposited to the \texttt{fData} slot.

We see the \texttt{fData} contains 25 columns describing information such as
the number of peptides, associated markers, machine learning results
etc. To identify the feature variable names we can use the function
\texttt{fvarLabels}. We see that the first 6 feature variable names contain
non-discriminatory label names, so we relabel them to help us identify
what feature data information is stored in the associated columns.

<<removefData>>=
fvarLabels(hl)
fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
fData(hl)[1:4, 1:6]
@

Note that when using the simple \texttt{readMSnSet2} procedure, the \texttt{pData}
slot which is used to store information about the samples/channels
is kept empty. It is advised to annotate the channels as
well. Below, we annotate the replicate from which the profiles
originate and the TMT tag (extracted from the sample/channel names).

<<pdata>>=
pData(hl)$Replicate <- rep(1:2, each = 10)
pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
pData(hl)
@

Throughout this workflow we refer to the different columns that
are found in the \texttt{exprs} (expression data) slot as channels (short for TMT channels).
In the frame of LOPIT and hyperLOPIT these channels consititue the
relative abundance of each protein (along the rows) in the channel of
interest. Each TMT channel originates from fractions collected from
the density gradient, or a set of pooled fractions or may be a sample
originating from an alternative preparation e.g. such as from the
chromatin enrichment performed in Christoforou et al \cite{hyper}.
Information about which gradient fractions were used for which tag
should also be stored in the sample meta-data \texttt{pData} slot.

The sample meta-data that is distributed with the
\Biocexptpkg{pRolocdata} package for Christoforou's
hyperLOPIT experiment and (as above) the quantitation data file, are
located in the \texttt{extdata} in the \Biocexptpkg{pRolocdata}
package on the hard drive.

In the code chunk below we again use the
\texttt{dir} function to locate the filepath to the meta-data \texttt{csv} file and
then read it into R using \texttt{read.csv}. We then append the meta-data to
the \texttt{pData} slot.  Information about the gradient fractions used and
the associated subcellular fraction densities in % w/v Iodixanol for
each replicate are stored here.

<<morepdata>>=
expinfo <- dir(extdatadir, full.names = TRUE,
               pattern = "hyperLOPIT-SIData-fraction-info.csv")

fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
                     header = FALSE, stringsAsFactors = FALSE)

pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
pData(hl)
@

\subsection*{Data processing}

\subsubsection*{Normalisation}

There are two aspects related to data normalisation that are relevant
to spatial proteomics data processing. The first one focuses on
reducing purely technical variation between channels without
affecting biological variability (i.e. the shape of the quantitative
profiles). This normalisation will depend on the underlying
quantitative technology and the experimental design, and will not be
addressed in this workflow. The second aspect, and more specific to
spatial proteomics data, is scaling all the organelle-specific
profiles into a same intensity interval (typically 0 and 1) by, for
example, diving each intensity by the sum of the intensities for that
quantitative feature. This is not necessary in this example as the
intensities for each replicate have already been re-scaled to 1 in
Proteome Discoverer. However, if one wanted to do this they would
execute the \texttt{normalise} function as demonstrated in the below code
chunk.

<<normsum, eval = FALSE>>= 
hl <- normalise(hl, method = "sum") 
@

This transformation of the data assures to cancel the effect of the
absolute intensities of the quantitative features along the rows, to
focus subsequent analyses on the relative profiles along the
sub-cellular channels.

The same \texttt{normalise} function (or \texttt{normalize}, both spellings are
supported) can also be applied in the first case described above.
Different normalisation methods such as mean or median scaling,
variance stabilisation or quantile normalisation, to cite a few, can
be applied to accomodation different needs.

As previously mentioned, before combination, the two replicates in the
\texttt{hl} data that we read into R were separately normalised by sum (i.e.
to 1) across the 10 channels for each replicate respectively. We can
verify this by summing each rows for each replicate:

<<normcheck>>=
summary(rowSums(exprs(hl[, hl$replicate == 1])))
summary(rowSums(exprs(hl[, hl$replicate == 2])))
@

We see that some features do not add up exactly to 1 due to rounding
errors after exporting to intermediate files. These small deviations
do not bear any consequences here.

\subsubsection*{Combining acquisitions}

The spreadsheet that was used to create the \texttt{hl} MSnSet
included the two replicates within one .csv file.  We also provide
individual replicates in the \Biocexptpkg{pRolocdata}
package. Below, we show how to combine \texttt{MSnSet} objects and,
subsequently, how to filter and handle missing values. We start by
loading the \Biocexptpkg{pRolocdata} package and the
equivalent replicates using the \texttt{data} function.

<<reps4combine>>=
library("pRolocdata")
data(hyperLOPIT2015ms3r1)
data(hyperLOPIT2015ms3r2)
@

At the R prompt, typing

<<pkgdata, eval=FALSE>>=
pRolocdata()
@

will list the \Sexpr{nrow(pRolocdata()$results)} datasets that are
available in \Biocexptpkg{pRolocdata}.

%% $

Combining data is performed with the \texttt{combine} function. This
function will inspect the feature and sample names to identify how to
combine the data. As we want our replicates to be combined along the
columns (same proteins, different sets of channels), we need to assure
that the respective sample names differ so they can be identified from
one another. The function \texttt{updateSampleNames} can be used do
this.

<<samplenames>>=
identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
sampleNames(hyperLOPIT2015ms3r1)
sampleNames(hyperLOPIT2015ms3r2)
@

In addition, to matching names, the content of the feature metadata
for identical feature annotations must match exactly across the data
to be combined. In particular for these data, we expect the same
proteins in each replicates to be annotated with the same UniProt
entry names and descriptions, but not with the same coverage of number
of peptides or peptide-spectrum matches (PSMs).

<<fvarnames>>=
fvarLabels(hyperLOPIT2015ms3r1)
fvarLabels(hyperLOPIT2015ms3r2)
@

Below, we update the replicate specific feature variable names and
remove the shared annotation.

<<fvarnames2>>=
fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
fvarLabels(hyperLOPIT2015ms3r2)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r2)[3:5], 2)
fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5] 
fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
fvarLabels(hyperLOPIT2015ms3r1)
fvarLabels(hyperLOPIT2015ms3r2)
@

We can now combine the two experiments into a single \texttt{MSnSet}:

<<combine, cache = cache>>=
combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
combined
@

More details above combining data are given in the dedicated
*Combining MSnSet instances* section of the \Biocpkg{MSnbase}
\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
  vignette}.

\subsubsection*{Missing data}

Missing data are a recurrent issue in mass spectrometry applications,
and should be addressed independently of this workflow
\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
described how a high content in missing values in spatial proteomics
data and their inappropriate handling leads to a reduction of
sub-cellular resolution. Missing data can be imputated using 
\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying the
imputation method is then determined by a \texttt{methods} parameter. In our
particular case, missing values are indicative of protein groups that
were not profiles in both replicates.

\begin{figure}[!ht]
  \centering
<<namap, out.width="\\textwidth">>=
image2(is.na(combined), col = c("black", "white"),
       main = "Missing values (white cells) after combining replicates")
@  
  \caption{Heatmap of missing values}
  \label{fig:namap}
\end{figure}

We prefer to remove proteins that were not assayed in both of our two
replicated experiments. This is done with the \texttt{filterNA} function that
removes features that contain more than a certain proportion (default
is 0) missing values.

<<filterNA>>=
combined <- filterNA(combined)
combined
@

When more than 2 data are to be combined and/or too many proteins were
not consistently assayed, leading to too many proteins being filtered
out, we suggest to implement an ensemble of classifiers voting on
protein-sub-cellular niche membership over the output of several
experiments (see section *Supervised machine learning* for the
description of sub-cellular assignments).

\section*{Quality Control}

Data quality is routinely examined through visualisation to verify
that sub-cellular niches have been separated along the gradient. Based
on De Duve's principle \cite{DeDuve:1981} proteins that co-localise
exhibit similar quantitation profiles across the gradient fractions
employed. One approach that has been widely used to visualise and
inspect high throughput mass spectrometry-based proteomics data is
principal components analysis (PCA). PCA is one of many dimensionality
reduction methods, that allow one to effectively summarise
multi-dimensional data in to 2 or 3 dimensions to enable
visualisation. Very generally, the original continuous
multi-dimensional data is transformed into a set of orthogonal
components ordered according to the amount of variability that they
describe. The \texttt{plot2D} method in \Biocpkg{pRoloc} allows one to
plot the principal components (PCs) of a dataset against one another,
by default the first two components are plotted on the x- and y-axis,
respectively (the \texttt{dims} argument can be used to plot other PCs). If
distinct clusters are observed, we assume that there is organellar
separation present in the data. Although, representing the
multi-dimensional data along a limited set of PCs does not give us a
hard quantitative measure of separation, it is extremely useful
summarising complex experimental information in one figure, to get an
simplified overview of the data.

In the code chunk below we produce a PCA plot of the mouse stem cell
dataset. One point on the plot represents one protein. We can indeed
see several distinct protein clusters. We specify \texttt{fcol = NULL}, which
means not to consider any feature variable to annotate the features
(proteins) with colours. We will see later how to use this to annotate
the PCA plot with prior information about sub-cellular localisation.

\begin{figure}[!ht]
  \centering
<<qcplot, message=FALSE, warning=FALSE, out.width="\\textwidth">>=
library("pRoloc")
plot2D(hl, fcol = NULL, col = "black")
@  
  \caption{PCA plot of the mouse stem cell data \texttt{hl}.}
  \label{fig:pcahl}
\end{figure}

In the first instance we advise one to visualise their data without
any annotation (i.e. with \texttt{fcol = NULL}), before proceeding with data
annotation. The identification of well resolved clusters in the data,
constitutes an unbiased assessment of the data structure,
demonstrating the successful separation of sub-cellular clusters.

**TODO** There are other dimensionality reduction methods available in
the \texttt{plot2D} function, which can be parametrised with the
\texttt{method} argument.  Advantages, disadvantages? See
\href{https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32}
very interesting gist with intriguing tSNE examples. (This might be
better to come after markers.)

It is also useful to visualise the relative intensities along the
gradient to identify channels displaying particularly low yield. This
can be done using the \texttt{plotDist} and \texttt{boxplot} functions, that plot
the protein profiles occupancy along the gradient (we also display the
mean channel intensities) and a \texttt{boxplot} of the column intensities.
In the two plots displayed below, we re-order the TMT channles to pair
corresponding channels in the two replicates (rather than ordering
the channels by replicate).

\begin{figure}[!ht]
  \centering
<<qcbx, out.width="\\textwidth">>=
par(mfrow = c(1, 2))
o <- order(hl$Iodixonal.Density)
plotDist(hl[, o], pcol = "#00000010")
lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
boxplot(exprs(hl[, o]), las = 2)
@  
  \caption{Protein profiles and distribution of channel intensities.}
  \label{fig:qcbx}
\end{figure}

\section*{Markers}

In the context of spatial proteomics, a marker protein is defined as a
well-known resident of a specific sub-cellular niche in a species
*and* condition of interest. Applying this to machine learning (ML),
and specifically supervised learning, for the task of protein
localisation prediction, these markers constitute the labelled training
data to use as input to a classification analyses. Defining well-known
residents, and obtaining labelled training data for ML analyses can be
time consuming, but it is important to define markers that are
representative of the multivariate data space and on which a
classifier will be trained and generated. \Biocpkg{pRoloc}
provides a convenience function, \texttt{addMarkers}, to directly add markers
to a \texttt{MSnSet} object, as demonstrated in the code chunk below. These
marker sets can be accessed using the \texttt{pRolocmarkers()}
function. Marker sets are stored as a simple named vector in R, and
originate from in-house user-defined sets of markers or from previous
published studies \cite{Gatto:2014b}, which are continuosly updated and 
integrated.

<<markers>>=
## List available marker sets
pRolocmarkers()
@

These markers can then be mapped to a \texttt{MSnSet}'s \texttt{featureNames}. The
mouse dataset used here has Uniprot IDs stored as the \texttt{featureNames}
(see \texttt{head(featureNames(hl))}) and the names of the vector of
the mouse markers stored in \Biocpkg{pRoloc} (\texttt{mmus} markers) 
are also Uniprot IDs (see \texttt{head(mrk)} in the code chunk below), so 
it is straightforward to match names between the markers and 
the \texttt{MSnSet} instance using the \texttt{addMarkers} function.

<<addmrkers>>=
## Use mouse markers
mrk <- pRolocmarkers(species = "mmus")
head(mrk)

## Add mouse markers
hl <- addMarkers(hl, mrk)
@

If the naming between the marker sets and the \texttt{MSnSet} dataset
are different, one will have to convert and match the proteins
according to the appropriate identifier. Sometimes, we find the
equivalent entry name, Uniprot ID or accession number is stored with
the data, which makes conversion between identifers relatively
straightforward. If this is not the case however, conversion can be
performed using \Biocpkg{biomaRt}, the Bioconductor
\href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation
  resouces} or any conversion softwares available online.

We now visualise these annotations along the PCA plot using the
\texttt{plot2D} function and then use the \texttt{addLegend} function to map the
marker classes to the pre-defined colours. We also display the data
along the first and seventh PCs using the \texttt{dims} argument. Note that
in this second call to the \texttt{plot2D} function, we have omitted the
\texttt{fcol} argument, as the \texttt{"markers"} feature variable name is the
default value. We choose to display PCs 1 and 7 to illustrate that
while upper principal components explain much less variability in the
data (2.23\% for PC7, as opposed to 48.41\% for PC1), we see that the
mitochondrial and peroxisome clusters can be differenciated, despite
the apparent overlap in the two first PCs.

\begin{figure}[!ht]
  \centering
<<plotmarkers, out.width="\\textwidth">>=
par(mfrow = c(1, 2))
plot2D(hl, main = "pRolocmarkers for mouse")
addLegend(hl, cex = .7)
plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
@  
  \caption{Annotated PCA plots of the \texttt{hl} data.}
  \label{fig:plotmarkers}
\end{figure}

The colours have been defined so as to enable to differenciate up to
30 classes. If more are provided, different character symbols
(circles, squares, ... and empty and solid symbols) are used. The
colours and the default plotting characters (solid dots for the
markers and empty circles for the features of unknown localisation)
can of course be changed, as described in the \texttt{setStockcol} manual
page.

As demonstrated in \cite{hyper} and illustrated in the PCA plot above,
the Golgi apparatus proteins display a dynamic pattern, noting sets of
Golgi marker proteins that are distributed amongst other subcellular
structures, an observation supported by microscopy. As such, we are
going to reset the annotation of Golgi markers to unknown using the
\texttt{fDataTounknown} function. It is often used to replace empty
strings ("") or missing values in the markers definition to a common
definition of *unknown localisation*.

<<removeGA>>=
hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
getMarkers(hl)
@

In general, the Gene Ontology (GO) \cite{Ashburner:2000}, and in
particular the cellular compartment (CC) namespace are a good starting
point for protein annotation and marker definition. It is important to
note however that automatic retrieval of sub-cellular localisation
information, from \Biocpkg{pRoloc} or elsewhere, is only the
beginning in defining a marker set for downstream analyses. Expert
curation is vital to check that any annotation added is in the correct
context for the the biological question under investigation.

Another useful visualisation that relies on marker annotation is the
representation of the protein profiles occupancy along the gradient
using the \texttt{plotDist} function. While the PCA plot enables to
efficiently visualise the complete dataset and assess the relative
separation of different sub-cellular niches, comparing profiles of a
few marker clusters is useful to assess how exactly they differ (in
terms of peak channels, for example). Below, we plot the profile of
the mitochondrial and peroxisome markers to highlight the differences
in profiles between these two sets of markers along the 6th and 7th
channels, as represented above along the 7th PC on the PCA plot.

\begin{figure}[!ht]
  \centering
<<plotDist, out.width="\\textwidth">>=
hlo <- hl[, order(hl$Iodixonal.Density)]
plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
         pcol = "purple", fractions = "Fraction.No")
title(main = "Marker occupancy profiles along the gradient")
matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
         lty = 1, col = "darkblue", type = "l")
legend("topleft", c("Mitochondrion", "Peroxisome"),
       lty = 1, col = c("purple", "blue"), bty = "n")
@  
  \caption{Mitochondrion and peroxisome protein profiles.}
  \label{fig:plotDist2}
\end{figure}

\textbf{TODO} Possibly mention \texttt{plot3D} and \texttt{mrkHClust}
(also mentioned later), although these are only available in the
development version.

\section*{Replication}

With the aim of maximising the sub-cellular resolution and,
consequently, the reliability in protein sub-cellular assignments, we
follow the advice in \cite{Trotter:2010} and combine replicated spatial
proteomics experiments as described above. Indeed, Trotter et
al. have shown a significant improvement in protein–organelle
association upon direct combination of single experiments, in
particular when these resolve different subcellular niches.

Direct comparisons of individual channels in replicated experiments
does not provide an adequate, goal-driven assessment of different
experiments. Indeed, due to the nature of the experiment and gradient fraction
collection, the quantitative channels 
do not correspond to identical selected fractions along the
gradient. As can be seen in the table below (taken from the
\texttt{hyperLOPIT}'s \texttt{pData}, focusing on channels 7 to 10), different sets
of gradient fractions are pooled to obtain enough material and optimise
acurate quantitation.

<<pdtab, echo=FALSE, eval = FALSE>>=
i <- c(2, 3, 4, 13, 14)
pData(hl)[i, ]
@

% On the figure below, we compare the relative intensities of channels 7
% to 10 on scatter plots, highlighting different marker sets. These
% differences result from different peak fractions for these niches on
% the mitochondrial and chromatin occupancy profiles for the two
% replicates.

% <<repl1, eval = TRUE, echo=FALSE, fig.width = 12>>=
% sel0 <- fData(hyperLOPIT)$markers != "unknown"
% cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
% cls[!sel0] <- getUnknowncol()

% par(mfrow = c(2, 4))
% for (i in c(7, 8, 10, 9)) {
%     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
%          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"]))
%     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
%     grid()
%     abline(0, 1, col = "red")
% }
% plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "")
% addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)

% i <- fData(hyperLOPIT)$markers == "Mitochondrion"
% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
%          featureNames(hyperLOPIT)[i], type = "l",
%          mcol = "orange")
% title(main = "Mitochondrion")
% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
% legend("topleft", paste("Replicate", 1:2),
%        col = c("#F781BF", "orange"),
%        lty = 1, bty = "n")

% i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
%          featureNames(hyperLOPIT)[i], type = "l")
% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
% title(main = "Chromatin")
% legend("topleft", paste("Replicate", 1:2),
%        col = c("#984EA3", "steelblue"),
%        lty = 1, bty = "n")
% @

The more relevant comparison unit is not a single channel, but rather
the complete protein occupancy profiles, which are best visualised as
experiment-wide on a PCA plot. As such, we prefer to focus on the
direct, qualitative comparison of individual replicate PCA plots,
assuring that each displays acceptable sub-cellular resolution. Note
that in the code chunk below, we mirror the x-axis to represent the
two figures with the same orientation.

\begin{figure}[!ht]
  \centering
<<plot2Drep, out.width="\\textwidth">>= 
par(mfrow = c(1, 2)) 
plot2D(hl[, hl$Replicate == 1], main = "Replicate 1") 
plot2D(hl[, hl$Replicate == 2], main = "Replicate 2", mirrorX = TRUE) 
@  
  \caption{PCA plots of replicates 1 and 2}
  \label{fig:plot2Drep}
\end{figure}

\section*{Interactive visualisation}

Visualisation and data exploration is an important aspect of data
analyses allowing one to shed light on data structure and patterns of
interest. Using the \Biocpkg{pRolocGUI} package we can
interactively visualise, explore and interrogate quantitative spatial
proteomics data. The \Biocpkg{pRolocGUI} package is currently
under active development and it relies on the \texttt{shiny} framework for
reactivity and interactivity. The package currently distributes 3
different GUI's ("main" (default), "compare" or "compare") which are
wrapped and launched by the \texttt{pRolocVis} function. In the below code
chunk we lauch the main app (note, we do not need to specify the
argument, \texttt{app = "main"} as it is the default).

<<mainapp, eval = FALSE, echo = TRUE>>=
library("pRolocGUI")
pRolocVis(hl)
@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./Figures/mainapp.png}
  \caption{A screen shot of clickable interface and zoomable PCA plot of the main app in the \Biocpkg{pRolocGUI} package.}
  \label{fig:app}
\end{figure}

As diplayed in the screenshot above, the "main" application is
designed for exploratory data analysis and is divied into 3 tabs: (1)
PCA, (2) Profiles and (3) Table selection. The default view upon
loading is the PCA tab, which features a clickable interface and
zoomable PCA plot with an interactive data table for displaying the
quantitation information. Particular proteins of interest can be
highlighted using the text search box. There is also an alternate
profiles tab for visualisation of the protein profiles, which can be
used to examine the patterns of proteins of interest. The Table
selection tab provides an interface to control data table column
selection.

The \textit{compare} application is useful for examining two replicate
experiments, or two experiments from different conditions, treatments
etc. The compare application is called by default if the input object
to \texttt{pRolocVis} is a \texttt{MSnSetList} of 2 \texttt{MSnSets},
but it can also be specified by calling the argument \texttt{app =
  "compare"}. For example, in the code chunk below we first create a
\texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT data, this
is then passed to \texttt{pRolocVis}.

<<compareapp, eval = FALSE, echo = TRUE>>=
data(hyperLOPIT2015ms3r1)
data(hyperLOPIT2015ms3r2)
mydata <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
pRolocVis(mydata, app = "compare")
@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./Figures/SS_Compare1.jpg}
  \caption{The compare application, main panel}
  \label{fig:compare}
\end{figure}

The comparison app loads the two PCA plots side-by-side. Only common
proteins between the two data sets are displayed. As per the main
application proteins can be searched, identified and highlighted on
both PCA plots and in the dedicated profiles tab. One key feature of
the compare application is the ability to re-map the second dataset
onto the PCA data space of the first (reference) data set (see
\texttt{?pRolocVis} and the argument \texttt{remap = TRUE}). Using the first dataset
as the reference set, PCA is carried out on the first dataset and the
standard deviations of the principal components (i.e. the square roots
of the eigenvalues of the covariance/correlation matrix) and the
matrix of variable loadings (i.e. a matrix whose columns contain the
eigenvectors) are stored and then used to calculate the principal
components of the second dataset. Both datasets are scaled and
centered in the usual way. The first dataset appears on the left, and
the second re-mapped data appears on the right. The order of the first
(the reference data for remapping) and second dataset can be changed
through regeneration/re-ordering of the \texttt{MSnSetList} object.

The final application "classify", has been designed to view machine
learning classification results according to user-specified thresholds
for the assignment of proteins to its sub-cellular location, this is
discussed later in the subsection \textit{thresholding} in the
\textit{supervised machine learning section}.

\section*{Novelty Detection}

The extraction of sub-cellular protein clusters can be difficult owing
to the limited number of marker proteins that exist in databases and
elsewhere. Furthermore, given the vast complexity of the cell,
automatic annotation retrieval does not always give a full
representation of the true sub-cellular diversity in the data. For
downstream analyses, such as supervised machine learning, it is
desirable to obtain reliable markers that cover as many sub-cellular
niches as possible, as these markers are directly used in the training
phase of the ML classification. We find that a lack of sub-cellular
diversity in the labelled training data leads to prediction errors, as
unlabelled instances can only be assigned to a class that exists in
the training data \cite{Breckels:2013}. In such scenarios novelty
detection can be useful to identify data-specific sub-cellular
groupings such as organelles and protein complexes. The phenotype
discovery (phenoDisco) algorithm \cite{Breckels:2013} is one such method
and is available in \Biocpkg{pRoloc}. It is an iterative
semi-supervised learning method that combines the classification of
proteins on existing labelled data with the detection of new clusters.

In addition to extracting new phenotypes, novelty detection methods
are also useful for confirming the presence of known or postulated
clusters in an unbiased fashion. For example, in \cite{hyper} the
\texttt{phenoDisco} algorithm was used to confirm the data-specific presence
of the nucleus and nucleus sub-compartments. In the code chunk below,
we demonstrate how to do this analysis, highlighting some of the
optional arguments and parameters available for phenotype extraction
and give some advice on how to interpret the output.

As the \texttt{phenoDisco} algorithm is semi-supervised it uses both labelled
(markers) and unlabelled data to explore the data structure and find
new sub-cellular data clusters. Thus the first step is to define some
input labelled data i.e. markers, that the algorithm will use as input
for the supervised learning aspect of the algorithm. As described in
\cite{hyper} we define a set of markers to use as input for the analyses
that cover well-known residents from three distinct organelle
structures; the mitochondria, plasma membrane and ER, and from three
well-known and abundant protein complexes; the proteasome and two
ribosomal subunits, 40S and 60S.  These input markers are stored in
the \texttt{featureData} column of \texttt{hl} where \texttt{fcol = "phenoDisco.Input"}. We
can use the convenience accessor function \texttt{getMarkers} to print out a
table of the markers contained in this marker set. These initial
markers were manually curated using information from the UniProt
database, the Gene Ontology and the literature.

<<displayPDmarkers>>=
getMarkers(hl, fcol = "phenoDisco.Input")
@

In the code chunk below we show how to run the \texttt{phenoDisco}
function and return a novelty detection result, according to the
specified parameters. The algorithm parameters \texttt{times}, \texttt{GS} and \texttt{p}
are passed to the function, along with the \texttt{fcol} to tell the
algorithm where the input training data is contained.

<<runPD, eval = FALSE>>=
## As per Christoforou et al (2016), 
hl <- phenoDisco(hl, fcol = "phenoDisco.Input", 
                 times = 200, 
                 GS = 60)
@

Note: We do not evaluate this code chunk in this document as the
algorithm is computational intensive and best parallelised over
multiple workers. This phenoDisco analysis took ~24 hours to complete
when parallelised over 40 workers.

The argument \texttt{times} indicated the number of times we run unsupervied
Gaussian Mixture Modelling before defining a new phenotype cluster.
The recommended minimum and default value is 100. In the above code
chunk we increase the value to \texttt{times = 200} as we have found for
larger datasets (e.g. 5000+ proteins) a higher times is requried for
convergence. \texttt{GS} defines the minimum number of proteins allowed per
new data cluster and thus heavily influences what type of new clusters
are extracted. For example, if a user is interesed in the detection of
small complexes they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
If they wish to detect larger more abundant sub-cellular niches a much
higher \texttt{GS} would be preferable. Specifying a small \texttt{GS} can be more
time consuming than using a larger \texttt{GS}, and there is a trade off
between finding interesting small complexes and those that may not be
of interest as we find there is a tendancy to find more noise when
using a small \texttt{GS} compared to using a higher one.

One may also consider increasing the search space for new data
clusters by increasing the value of the parameter \texttt{G}. This defines
the number of GMM components to test and fit; the default is \texttt{G = 1:9}
(\texttt{mclust}'s default). One should note that the decreasing the \texttt{GS},
and increasing the values of the arguments \texttt{times}, \texttt{G} (among other
function arguments, see \texttt{?phenoDisco}) will heavily influence
(increase) the total time taken to run the algorithm.  \texttt{phenoDisco}
supports parallelisation and we strongly suggest you make use of a
parallel processing to run these analyses (different backends can be
set with argument \texttt{BPPARAM}).

<<loadPDres, echo=FALSE>>=
f0 <- dir(extdatadir, full.names = TRUE,
          pattern = "bpw-pdres.rds")
pdres <- readRDS(f0)
hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
setStockcol(NULL)
setStockcol(paste0(getStockcol(), 70))
@

The ouput of running the \texttt{phenoDisco} algorithm is a \texttt{MSnSet}
containing the new data clusters, appended to the \texttt{featureData} under
the name \texttt{pd}. We can see by typing \texttt{hl@processingData} directly into
the console the processing information has been updated to the
\texttt{MSnSet} recording the parameters that were used to run the analyses.
This is handy for keeping track of data analyses. The results can be
displayed by using the \texttt{getMarkers} function. We see that 5 new
phenotype data clusters were found.

<<showPDres>>=
hl
getMarkers(hl, fcol = "pd")
@

We can plot the results using the \texttt{plot2D} function.

\begin{figure}[!ht]
  \centering
<<plotPDres, out.width="\\textwidth">>=
## Re-order the colours for the phenoDisco output
cl <- getMarkerClasses(hl, "pd")
cols <- getStockcol()[seq(cl)]
ind <- grep("Pheno", cl, invert = TRUE)
cols[ind] <- getStockcol()[seq(cl)][1:length(ind)]
cols[-ind] <- getStockcol()[seq(cl)][(length(ind) + 1):length(cl)]

## Plot the input and output
par(mfrow = c(1, 2))
plot2D(hl, fcol = "phenoDisco.Input", 
       main = "phenoDisco input markers", col = getStockcol()[1:6])
addLegend(hl, fcol = "phenoDisco.Input", cex = .7)
plot2D(hl, fcol = "pd", main = "phenoDisco output", col = cols)
addLegend(hl, fcol = "pd", cex = .7, col = cols)
@
  \caption{Results of the novelty detection algorithm.}
  \label{fig:plotPDres}
\end{figure}

The five new phenotype data clusters can be extracted and examined.
In the code chunk below we write the results to a .csv file. We use
the argument \texttt{fDataCols} to specify which columns of the
\texttt{featureData} to write.

<<writehl, eval=FALSE>>=
fData(hl)$pd <- as.character(fData(hl)$pd)
write.exprs(hl, fDataCols = "pd", file = "pd-results.csv", sep = ",")
@

We can also examine the each phenotype intercatively and visualise
their protein profiles by using the \texttt{pRolocVis} function in the
\Biocpkg{pRolocGUI} package. We found that phenotype 1 was
enriched in nucleus associated proteins, phenotype 2 in chromatin
associated proteins, phenotype 3 in cytosolic and phenotypes 4 and 5
in lysosomal and endosomal proteins.

<<vishl, eval=FALSE>>=
pRolocVis(hl, fcol = "pd")
@

\section*{Supervised machine learning}

Supervised machine learning, also known as classification, is an
essential tool for the assignment of proteins to distinct sub-cellular
niches. Using a set of labelled training examples i.e. markers, we can
train a machine learning classifier to learn a mapping between the
data i.e. the quantitative protein profiles, and a known localisation.
The trained classifier can then be used to predict the localisation of
a protein of unknown localisation, based on its observed protein
profile. To date, this method has been extensively used in spatial
quantitative proteomics to assign thousands of proteins to distinct
sub-cellular niches
\cite{hyper,Groen:2014,Trotter:2010,Hall:2009,Dunkley:2006,Tan:2009}.

There are several classification algorithms readily available in
\Biocpkg{pRoloc}, which are documented in the dedicated
\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\Biocpkg{pRoloc}
  machine learning techniques vignette}.  We find the general tendancy
to be that it is not the choice of classifier, but the improper
optimisation of the algorithmic parameters, that limits classification
accuracy. Before employing any classification algorithm and generating
a model on the training data, one must first find the optimal
parameters for the algorithm of choice.

\subsection*{Optimisation}

In the code chunk below we employ the use of a Support Vector Machine
(SVM) to learn a classifier on the labelled training data. As
previously mentioned one first needs to train the classifiers
parameters before an algorithm can be used to predict the class labels
of the proteins with unknown location. One of the most common ways to
optimise the parameters of a classifier is to partition the labelled
data in to training and testing subsets. In this framework parameters
are tested via a grid search using cross-validation on the training
partition. The best parameters chosen from the cross-validation stage
are then used to build a classifier to predict the class labels of the
protein profiles on the test partition. Observed and expected
classication results can be compared, and then used to assess how well
a given model works by getting an estimate of the classiers ability to
achieve a good generalisation i.e. that is given an unknown example
predict its class label with high accuracy. In
\Biocpkg{pRoloc}, algorithmic performance is estimated using
stratified 80/20 partitioning for the training/testing subsets
respectively, in conjuction with five-fold cross-validation in order
to optimise the free parameters via a grid search. This procedure is
usually repeated 100 times and then the best parameter(s) are selected
upon investigation of classifier accuracy, here we use the harmonic
mean of precision and recall; the macro F1 score. In the code chunk
below we demonstrate how to optimise the free parameters; \texttt{sigma} and
\texttt{cost}, of a classical SVM classifier with a Gaussian kernel using the
function \texttt{svmOptimisation}. As the number of labelled instances per
class varies from organelle to organelle, we can account for class
imbalance by setting specific class weights when generating the SVM
model. Below the weights, \texttt{w} are set to be inversely proportional to
the class frequencies.

<<loadpdRes, echo=FALSE>>=
svmf <- dir(extdatadir, full.names = TRUE,
              pattern = "bpw-svmopt.rds") 
params <- readRDS(svmf)
@

<<getsvmweight>>=
w <- table(getMarkers(hl, verbose = TRUE))
w <- 1/w[names(w) != "unknown"]
@

<<dosvmopt, eval=FALSE>>=
## 100 rounds of optimisation with five-fold cross-validation
params <- svmOptimisation(hl, fcol = "markers",
                          times = 100, xval = 5,
                          class.weights = w)
@

As mentioned previously, we reply on the default feature variable
\texttt{"markers"} to define the class labels and hence can ommit
it. To use another feature variables, one need to explicitly specify
its name using the \texttt{fcol} argument (for example \texttt{fcol =
  "markers2"}).

The output \texttt{params} is an object of class \texttt{GenRegRes}; a
dedicated container for the storage of the design and results from a
machine learning optimisation. To assess classifier performance we can
examine the macro F1 scores and the most frequently chosen
parameters. A high macro F1 score indicates that the marker proteins
in the test dataset are consistently and correctly assigned by the the
algorithm. Often more than one parameter or set of parameters gives
rise to the best generalisation accuracy. As such it is always
important to investigate the model parameters and critically assess
the best choice. The best choice may not be as simple as the parameter
set that gives rise to the highest macro F1 score and one must be
careful to avoid overfitting and to choose parameters wisely.

<<visualiseOpt, eval=FALSE>>=
(best <- getParams(params))
plot(params)
levelPlot(params)
@

\begin{figure}[!ht]
  \centering
<<visualiseOptHide, echo=FALSE, out.width="\\textwidth">>=
library('gridExtra')
grid.arrange(plot(params), levelPlot(params), ncol = 2)
@  
  \caption{Assessment of the classification model parameter optimisation.}
  \label{fig:visualisOptHide}
\end{figure}

By using the function \texttt{getParams} we can extract the best set of
parameters. Currently, \texttt{getParams} retrieves the first best set is
automatically but users are encouraged to critically assess whether
this is the most wise choice by visualising the results using the
methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method for \texttt{GenRegRes}
object shows the respective distributions of the 100 macro F1 scores
for the best cost/sigma parameter pairs, and \texttt{levelPlot} shows the
averaged macro F1 scores, for the full range of parameter values. Once
we have selected the best parameters we can then use them to build a
classifier from the labelled marker proteins.

\subsection*{Classification}

We can use the function \texttt{svmClassification} to return a classification
result for all unlabelled instances in the dataset corresponding to
their most likely sub-cellular location. The algorithm parameters are
passed to the function, along with the class weights. As above, \texttt{fcol}
can be ignored as we use the labels defined in the default \texttt{"markers"}
feature variable.

<<classify>>=
hl <- svmClassification(hl, params,
                        class.weights = w,
                        fcol = "markers")
@

Automatically, the output of the above classification; the organelle
predictions and assignment scores, are stored in the
\texttt{featureData} slot of the \texttt{MSnSet}. In this case, they
are given the labels \texttt{svm} and \texttt{svm.scores} for the
predictions and scores respectively. The resultant predictions can be
visualised using \texttt{plot2D}. In the code chunk below
\texttt{plot2D} is called to generate a PCA plot of the data and
\texttt{fcol} is used to specify where the new assignments are located
e.g.  \texttt{fcol = "svm"}.

Additionally, when calling \texttt{plot2D} we can use the \texttt{cex}
argument to change the size of each point on the plot (where one point
represents one protein) to be inversely proportional to the SVM
score. This gives an initial overview of the high scoring
localisations from the SVM predictions.

<<setColors, include=FALSE>>=
setStockcol(NULL)
setStockcol(paste0(getStockcol(), 70))
par(mfrow = c(1, 1))
@

\begin{figure}[!ht]
  \centering
<<plotSVM, out.width="\\textwidth">>=
## set point size of each protein to be inversely proportional to the
ptsze <- exp(fData(hl)$svm.scores) - 1
## plot new predictions
plot2D(hl, fcol = "svm", cex = ptsze)
addLegend(hl, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
@  
  \caption{Classification results.}
  \label{fig:plotSVM}
\end{figure}

The adjustment of the point size intuitively confers important
information that is more difficult to defined formally, but that we
will address in the next section. The classifier (SVM in our case, but
this is also valid of other classifiers) defines boundaries based on
the labelled marker proteins. These class/organelle boundaries define
how non-assigned proteins are classified and with what confidence.

\textbf{TODO} explain how choosing markers positively/negatively
affects this, and relation with thresholding.

\subsubsection*{Thresholding}\label{sec:thresholding}

It is common when applying a supervised classification algorithm to
set a specific score cutoff on which to define new assignments, below
which classifications are set to unknown/unassigned. This is important
as in a supervised learning setup proteins can only be predicted to be
localised to one of the sub-cellular niches that appear in the
labelled training data. We can not guaranette (and do not expect) that
the whole class diversity to be represented in the labelled training
data as (1) finding markers that represent the whole diversity of the
cell is challenging (especially obtaining dual- and multiply-localised
protein markers) and (2) many sub-cellular niches contain too few
proteins to train on (we recommend a minimum of 13 markers per
sub-cellular class for stratified 80/20 partitioning and 5-fold
cross-validation - this allows a minimum of ~10 examples for parameter
optimisation on the training partition i.e. 2 per fold for 5-fold
cross-validation, and then ~3 for testing the best parameters on the
validation set). <!--Conversely, when defining markers one must also
be careful to not label too many markers, as this can lead to
overfitting and mis-assignment.-->

Deciding on a threshold is not trivial as classifier scores are
heavily dependent upon the classifier used and different sub-cellular
niches can exhibit different score distributions, as highlighted in
the boxplot below. We recommend users to set class-specific
thresholds.  In the code chunk below we display a boxplot of the score
distributions per organelle.

\begin{figure}[!ht]
  \centering
<<thresholds, out.width="\\textwidth">>=
## First remove the markers
preds <- unknownMSnSet(hl)
## Plot a boxplot of the scores of each organelle
par(oma = c(10.5, 0, 0, 0))
boxplot(svm.scores ~ svm, data = fData(preds), 
        ylab = "SVM scores", las = 2)
@
  \caption{Visualistion of class-specific classification score distribution.}
  \label{fig:threshold}
\end{figure}

There are many ways to set thresholds and the choice of method will
depend on the biological question and experimental design at hand. One
viable approach in the frame of the above experimetal design would be
to manually set a FDR, say 5\%, per organelle. To do this the user
would examine the top scoring predictions for each organelle, and then
set a threshold at the score at which they achieve 5\% of false
assignments per organelle.  The definintion of a false assignment
would depend on the information available, for example, validity or
lack of validity for the localisation from another experiment as
reported in the literature or a reliable database.  If such
information is not available, one crude method is to set a threshold
per organelle by extracting the median or 3rd quantile score per
organelle.  For example, in the code chunk below, we use the
\texttt{orgQuants} function to extract the median organelle scores and
then pass these scores to the \texttt{getPrediction} function to
extract the new localisations that meet this scoring criteria. Any
sub-cellular predictions that fall below the specified thresholds are
labelled as unknown.

<<orgquants>>=
ts <- orgQuants(hl, 
                fcol = "svm", 
                scol = "svm.scores",
                mcol = "markers", 
                t = .5)

hl <- getPredictions(hl,
                     fcol = "svm",
                     scol = "svm.scores", 
                     mcol = "markers",
                     t = ts)                        
@

The output of \texttt{getPredictons} is the original \texttt{MSnSet}
dataset with a new feature variable appended to the feature data
called \texttt{fcol.pred} (i.e. in our case \texttt{svm.pred})
containing the prediction results. The results can also be visualied
using \texttt{plot2D} function.

<<resetpar>>=
dev.off()
@

\begin{figure}[!ht]
    \centering
<<plotres, out.width="\\textwidth">>=
plot2D(hl, fcol = "svm.pred")
@  
  \caption{Results of the localisation preductions after thresholding.}
  \label{fig:plotres}
\end{figure}

There is also a dedicated interactive application to help users
examine these distributions in the \Biocpkg{pRolocGUI} package.
This app can be launched via the \texttt{pRolocVis} function and specifying
the argument \texttt{app = "classify"} along with the relevent \texttt{fcol}, \texttt{scol}
and \texttt{mcol} which refer to the columns in the feature data that contain
the new assignments, assignment scores and markers respectively (see
also \texttt{fvarLabels(svmres)}).

<<classifyApp, eval=FALSE>>=
library("pRolocGUI")
pRolocVis(hl, 
          app = "classify",
          fcol = "svm",
	        scol = "svm.scores",
          mcol = "markers")
@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./Figures/classify.png}
  \caption{The classify application}
  \label{fig:classifyapp}
\end{figure}

The data is loaded and displayed on a PCA plot and a boxplot is used
to display the classifier scores by data class. On the left there is a
sidebar panel with sliders to control the thresholds upon which
classifications are made. There are two types of cut-off that the user
can choose from: (1) "Quantile" and (2) "User-defined". By default,
when the application is launched quatile scoring is selected and set
to 0.5, the median. The class-specific score thresholds that
correspond to selecting the desired quantile are shown on as red dots
on the boxplot. The assignments on the PCA plot are also updated
according to the selected threshold. The quantile threshold can be set
by moving the corresponding quantile slider. If one wished to set
their own cut-offs the "User-defined" radio button must be selected
and then the sliders for defining user-specified scores become active
and the scores and highlighted on the boxplot by blue dots. For more
information we refer users to the \Biocpkg{pRolocGUI}
tutorial
\href{http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html}{vignette}.

\subsection*{Transfer learning}

In addition to high quality MS-based quantitative proteomics data
there exist a number of other sources of information that are freely
available in the public domain that may be useful to assign a protein
to its sub-cellular niche. For example, imaging from
immunofluorescence microscopy, protein annotations and sequences, and
protein-protein interactions among others, represent a rich and vast
source of complementary information. We can integrate this auxiliary
information with our primary MS-based quantitative data using a
paradigm known as transfer learning (TL). The integration of data
between different technologies is one of the biggest challenges in
computational biology to date and the \Biocpkg{pRoloc} package
provides functionality to do such analyses. We recently developed two
transfer learning algorithms using a *k*-NN and SVM framework and
applied them to the task of protein localisation prediction
\cite{Breckels:2016}. In this section we will begin with explaining the
concept of transfer learning and then show how to apply this in the
frame of spatial proteomics and protein localisation prediction.

In TL one typically has a primary task that they wish to solve, and
some complementary (often heterogeneous) auxiliary information that is
related to the primary learning objective, that can be used to help
solve the primary goal. For example, here our primary task is to
assign proteins to their sub-cellular niche with high generalisation
accuracy from data collected from quantitative MS-based experiments.
In the example below we extract Gene Ontology (GO) information to use
as an auxiliary data source to help solve our task of protein
localisation prediction.

Using the functions \texttt{setAnnotationParams} and
\texttt{makeGoSet} we can contruct an auxiliary \texttt{MSnSet} of GO
terms, from the primary data's features i.e. the protein accession
numbers. All the GO terms associated to each accession number are
retrieved and used to create a binary matrix where a 1 (0) at position
(i; j) indicates that term j has (not) been used to annotate protein
i. The GO terms are retrieved from an appropriate repository using the
\Biocpkg{biomaRt} package. The specific Biomart repository and query
will depend on the species under study and the type of
identifiers. The first step is to construct the annotation parameters
that will enable to perform the query, this is done using
\texttt{setAnnotationParams}. Typing into the R console \texttt{par <-
  setAnnotationParams()} will present two menus, firstly asking you to
identify the species of study, and then what type of identifier you
have used to annotate the proteins in your \texttt{MSnSet}. It is also
possible to pass patterns to match the species e.g. in the code chunk
below we pass "Mus musculus", and the identifier type for our data
(see \texttt{featureNames(hl)}) which is "Uniprot/Swissprot", for the
Biomart query.

<<setap, eval=FALSE>>=
par <- setAnnotationParams(inputs = c("Mus musculus", 
                                      "UniProt/Swissprot"))

@

Now we have contructed the query parameters we can use the
\texttt{makeGoSet} function to retrieve and build an auxiliary GO
\texttt{MSnSet} as described above. By default, the cellular component
terms are downloaded, without any filtering on evidence codes. It is
also possible to download terms from the molecular function and
biological process GO namespaces, and also apply filtering based on
evidence codes as desired, see \texttt{?makeGoSet} for more details.

<<makego, eval=FALSE>>=
gocc <- makeGoSet(hl, params = par, 
                  namespace = "cellular_component",
                  evidence = NULL)
@

The function \texttt{makeGoSet} uses the \Biocpkg{biomaRt} package to
query the relevent database (e.g. Ensembl, Uniprot) for GO terms. All
GO terms that have been observed for the 5032 proteins in the
hyperLOPIT dataset are retieved. Users should note that the number of
GO terms retreived is also dependent on the database version queried
and thus is always subject to change. We find it is common to see many
GO terms with only one protein assigned to that term, these such terms
generally do not bring and information generally for building the
classifier and thus we can remove such GO terms by using the function
\texttt{filterBinMSnSet}.

<<filtergo, eval=FALSE>>=
gocc <- filterBinMSnSet(hl)
@

Now we have generated our auxiliary data we can use the \textit{k}-NN
implementation of transfer learning available in \Biocpkg{pRoloc} to
integrate this with our primary MS-based quantitative proteomics data
using the functions \texttt{knntlOptimisation} to estimate the
free-parameters for the integration, and \texttt{knntlClassification}
to do the predictions. We have shown that using transfer learning in
the context of spatial proteomics results in the assignment of
proteins to sub-cellular niches with a higher generalisation accuracy
than using standard supervised machine learning with a single source
of information \cite{Breckels:2016}.

The first step, as with any machine learning algorithm, is to optimise
any free paramaters of the classifier. For the *k*-NN TL classifier
there are two sets of parameters that need optimising: the first set
are the $k$'s for the primary and auxiliary data sources required for
the nearest neighbour calculations for each data source. The second
set of parameters (which is noted by a vector of $\theta$ weights)
that require optimising are the class weights, one per subcellular
niche, that control the proportion of primary and auxiliary data to
use for learning. A weight can take any real value number between 0
and 1. A weight of $\theta = 1$ indicates that all weight is given to
the primary data (and this implicitly implies that a weight of
$1 - \theta$ is given to the auxiliary data, similarly a weight of
$\theta = 0$ implies that all weight is given to the auxiliary data
(so 0 is given to the primary source). If we conduct a parameter seach
and test weights $\theta = {0, 1/3, 2/3, 1}$ for each class, and if we
have, for example 10 subcellular niches, this will result in
\texttt{4$^{10}$} different combinations of parameters to test. The
parameter optimisation is therefore time consuming and as such we
recommend you make use of a computing cluster (code and submissing
scripts are also available in the supporting information). The markers
in the \texttt{hl} dataset contain 14 subcellular classes. If we
examine these markers and classes on the PCA plot above we can see
that in particular the two ribosomes and two nuclear compartments are
highly separated along the first two components, this is also evident
from the profiles plot which gives us a good indication that these
subcellular niches are well-resolved in the hyperLOPIT
dataset. Transfer learning is particularly useful for classes that are
not as well separated, we find that subcellular niches that are
well-separated under hyperLOPIT and LOPIT obtain a class score of 1
(i.e. use only primary data from transfer learning
\cite{Breckels:2016}). Therefore, for the optimisation stage of the
analyses we can already infer a subcellular class weight of 1 for
these niches and only optimise over the remaining organelles.  This
can signifciantly cut down optimisation time as by removing these 4
classes from the optimisation (and not the classification) we only
have \texttt{4$^{10}$} class weight combinations to consider instead
of \texttt{4$^{14}$} combinations.

In the example below we remove these 4 classes from the marker set,
re-run the \texttt{knnOptimisation} for each data source and then run
the \texttt{knntlOptimisation} with the 10 remaining classes. (Note:
this is not run live as this the \texttt{hl} dataset with 10 classes,
707 markers and 4$^{10}$ combinations of parameters takes ~76 hours to
run on the University of Cambridge HPC using 256 workers).

To remove the 4 classes and create a new column of markers in the
feature data called \texttt{tlmarkers} to use for the analysis:

<<loadgocc, include=FALSE>>=
gofile <- dir(extdatadir, full.names = TRUE,
              pattern = "bpw-gocc.rds")
gocc <- readRDS(gofile)
@

<<setuptlopt>>=
## create new markers column for tl markers
fData(hl)$tlmarkers <- fData(hl)$markers
fData(gocc)$tlmarkers <- fData(gocc)$markers

## Remove 4 classes
torm  <- c("40S Ribosome", "60S Ribosome",
           "Nucleus - Chromatin", 
           "Nucleus - Non-chromatin")
for (i in seq(torm)) {
  hl <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
  gocc <- fDataToUnknown(gocc, from = torm[i], fcol = "tlmarkers")
}
getMarkerClasses(hl, fcol = "tlmarkers")
getMarkerClasses(gocc, fcol = "tlmarkers")
@

Optimisation stage 1: run \texttt{knnOptimisation} to get the best
$k$'s for each data source.

<<knnoptfortl, eval=FALSE>>=
## get best k's
kpopt <- knnOptimisation(hl, fcol = "tlmarkers")
kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
plot(kpopt)
plot(kaopt)
@

From examining the parameter seach plots we find the best $k$'s for
both the primary and auxiliary are 3.

Optimisation stage 2: run \texttt{knntlOptimisation} to get the best
transfer learning weights for each sub-cellular class.

<<runtlopts, eval=FALSE>>=
## Set appropriate parallelisation backend and  
## number of workers for the tl
par <- SnowParam(255L, type = "MPI")

## Now peform tl optimisation 
tlopt <- knntlOptimisation(hl, gocc,
                           fcol = "tlmarkers",
                           length.out = 4,
                           times = 50,
                           xval = 5, k = c(3, 3),
                           BPPARAM = par)
@

The results of the optimisation can be visalised using the \texttt{plot}
method for \texttt{"ThetaRegRes"} objects:

<<loadtlres, echo=FALSE>>=
tlfile <- dir(extdatadir, full.names = TRUE,
              pattern = "bpw-tlopt.rds")
tlopt <- readRDS(tlfile)
colnames(tlopt@results)[4] <- "Endoplasmic reticulum"
setStockcol(NULL)
@

\begin{figure}[!ht]
  \centering
<<plottl, out.width="\\textwidth">>=
plot(tlopt)
@  
  \caption{Visualisation of the tTransfer learning parameter optimisation procedure.}
  \label{fig:plottl}
\end{figure}

Looking at the bubble plot displaying the distribution of best weights
over the 50 runs we find that for many of the subcellular niches a
weight of 1 is most popular (i.e. use only primary hyperLOPIT data in
classification), this is unsuprising as we already know the dataset is
well resolved for these classes. We see that the most popular weights
for the proteasome and lysosome tend to be towards 0, indicating that
these niches are well-resolved in the Gene Ontology. This tells us
that we would benefit from including auxiliary GO information in our
classifier for these subcellular compartments. The plasma membrane
weights are relatively equally spread between using hyperLOPIT and GO
data. Using the \texttt{getParams} function we can return the best
weights and then use this as input for the classification.

One of the benefits of the algorithm is the ability to manually select
weights for each class. In the optimisation above, for time
constraints, we removed the two ribosomal subunits and the two nulcear
compartments, and therefore in the code chunk below when we extract
the best parameters, these subcellular niches are not included. To
include these 4 subcellular niches in the next classification step we
must include them in the parameters so we infer a weight of 1 for each
of these niches as we know they are well resolved in hyperLOPIT. We
then re-order the weights according to \texttt{getMarkerClasses} and
perform the classification using the function
\texttt{knntlClassification}.

<<tlclassify, cache = cache>>=
## best parameters for the 10 classes
(bestpar <- getParams(tlopt))

## add weights for classes not included in the optimisation
otherweights <- rep(1, 4)
names(otherweights) <- c("40S Ribosome", "60S Ribosome", 
                         "Nucleus - Chromatin", 
                         "Nucleus - Non-chromatin")
(bestpar <- c(bestpar, otherweights))

## re-order classes 
bestpar <- bestpar[getMarkerClasses(hl)]

## Do the classification
hl <- knntlClassification(hl, gocc, bestTheta = bestpar, 
                          k = c(3, 3))
@

The results from the classification results and associated scores are
appended to the \texttt{fData} slot and named \texttt{knntl} and
\texttt{knntl.scores} respectively. Results can be visualised using
\texttt{plot2D}, scores assessed and cutoffs calculated using the
\texttt{classify} app in \texttt{pRolocVis}, predictions obtained
using \texttt{getPredictions} in the same way as demonstrated above
for the SVM classifier.

\subsection*{Unsupervised machine learning}

In \texttt{pRoloc} there is functionality for unsupervsied machine
learning methods. In unsupervised learning, the training data consists
of a set of input vectors e.g. protein profiles, for which we do not
have information about the class label e.g. localisation. The main
goal in unsupervised learning is to uncover groups of similar examples
within the data, this is termed clustering. Ordinance methods such as
principal components analysis (PCA) also fall into the category of
unsupervised learning methods, where the data can be projected from a
high-dimensional spcae down to two or three dimensions for the purpose
of visualisation.

As described and demonstrated already above, PCA is a valuable and
powerful method for data visualisation and quality control. We do not
do any unsupervised clustering in the frame of the above spatial
proteomics workflow as although clustering can be useful for grouping
labelled data into categories e.g. see the \texttt{mrkHClust}
function, we do not find it adaquete for spatial proteomics data
analysis. We find supervised learning more suited to the task of
protein localisation prediction in which we use high-quality curated
marker proteins to build a classifier, instead of using an entirely
unsupervised approach to look for clusters and then look for
enrichment of organelles and complexes. In the latter we do not make
good use of valuable prior knowledge, and in our experience
unsupervised clustering can be extremely difficult due poor estimates
of the number of clusters that may appear in the data.

\section*{Writing and exporting data}

A \texttt{MSnSet} can be exported from R using the
\texttt{write.exprs} function.  This function writes the expression
values to a tab (the defualt, or other type e.g.e.g. comma, as
specified in \texttt{write,table}) separated file. There argument
\texttt{fDataCols} can be used to specify which \texttt{featureData}
columns (as column names, column number or \texttt{logical}) to append
to the right of the expression matrix.

In the below code chunk we write the \texttt{hl} object to a csv
file. The \texttt{file} argument is used to specify the file path, the
\texttt{sep} argument specifies the field separator string, here we
use a comma, finally as we want to write all the information in the
\texttt{featureData} to the file, as well as the expression data, we
specify \texttt{fDataCols = 1:ncol(hl)} i.e. write everything in the
\texttt{featureData} to the file \texttt{"hl.csv"}.

<<writeData, eval = FALSE>>=
write.exprs(hl, file = "hl.csv", sep = ",", 
            fDataCols = 1:ncol(hl))
@

Individual R objects, such a parameters from the machine learning
optimisations, can be save using the standard \texttt{save} function
in R.  For example, to save and then re-load the parameters from the
SVM optimisation,

<<saveData, eval = FALSE>>=
## To save the parameters as an R object
save(params, file = "svmparams.rda")

## To re-load after saving
load(file = "svmparams.rda")
@

\section*{Session information}

The function \texttt{sessionInfo} provides a summary of all packages
and versions used to generate this document. This enables us to record
the exact state of our session that lead to these exact
results. Conversely, if the script stops working of if it returns
different results, we are in a position to re-generate the original
results using the adequate software versions and retrace changes in
the software that lead to failure and/or different results.

<<si>>=
sessionInfo()
@

It is always important to include session information details along
with a \href{http://adv-r.had.co.nz/Reproducibility.html}{short
  reproducible example} highlighting the problem or
\href{https://support.bioconductor.org/}{question} at hand.