### 2018-04-03 main.tex
add f1000 files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Welcome to Overleaf --- just edit your article on the left,
% and we'll compile it for you on the right. If you give 
% someone the link to this page, they can edit at the same
% time. See the help menu above for more info. Enjoy!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% For more detailed article preparation guidelines, please see:
% http://f1000research.com/author-guidelines

%% \documentclass[10pt,a4paper,twocolumn]{article}
\documentclass[10pt,a4paper]{article}
\usepackage{f1000_styles}

%% Default: numerical citations
\usepackage[numbers]{natbib}

%% Uncomment this lines for superscript citations insteadl
% \usepackage[super]{natbib}

%% Uncomment these lines for author-year citations instead
% \usepackage[round]{natbib}
% \let\cite\citep

\usepackage{hyperref}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{knitr}
\usepackage{alltt}

\definecolor{fgcolor}{rgb}{.5,.6,.7}

\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}

\newcommand{\Rpackage}[1]{\textsl{#1}}
\newcommand\Biocpkg[1]{%
  {\href{http://bioconductor.org/packages/#1}%
    {\Rpackage{#1}}}}
\newcommand\Biocannopkg[1]{\Biocpkg{#1}}
\newcommand\Biocexptpkg[1]{\Biocpkg{#1}}

\begin{document}

\title{A Bioconductor workflow for processing and analysing spatial proteomics data}

% \titlenote{The title should be detailed enough for someone to know
% whether the article would be of interest to them, but also
% concise. Please ensure the broadness and claims within the title are
% appropriate to the content of the article itself.}

\author[1,2]{Lisa M. Breckels}
\author[2]{Claire M. Mulvey}
\author[2]{Kathryn S. Lilley}
\author[1,2]{Laurent Gatto}
\affil[1]{Computational Proteomics Unit, Cambridge Systems Biology Centre, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}
\affil[2]{Cambridge Centre for Proteomics, Department of Biochemistry, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}

\maketitle
\thispagestyle{fancy}

\begin{abstract}
\input{abstract}
\end{abstract}

\input{bioc-workflow}

\bigskip

{\small\bibliographystyle{unsrtnat}
\bibliography{bibliography}}

\end{document}